---
title: "GUV_Analysis_summary"
date: "`r Sys.Date()`"
author: "James Foster"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
editor_options: 
  markdown: 
    wrap: sentence
---

## Details

> **AUTHOR**
>
> James Foster $2025~06~05$

> **MODIFIED**
>
> James Foster $2025~06~28$

> **DESCRIPTION**
>
> Loads the reorganised data from `colour_dance_reorg.csv`, saved from `GUV_organisedata.R`, and generates plots to show the how dances changed as a function of light stimulus.

> **INPUTS**
>
> `colour_dance_reorg.csv`, `Edrich1979_brightness-accuracy.csv`

> **OUTPUTS**
>
> Plotted results.

> **CHANGES**
>
> -   

> **REFERENCES**
>
> Edrich, W., Neumeyer, C.
> and von Helversen, O.
> (1979).
> “Anti-sun orientation” of bees with regard to a field of ultraviolet light.
> *J. Comp. Physiol.* 134, 151–157.
>
> Rossel, S.
> and Wehner, R.
> (1984).
> Celestial orientation in bees: the use of spectral cues.
> *J. Comp. Physiol. A* 155, 605–613.

> **TODO**
>
> -   move necessary functions to GUV_functions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up

First load the functions and packages required for this project.
These can all be found in `GUV_functions.R`, within the same Github repository.

```{r Load functions and packages}
source('GUV_functions.R')
```

Then load the data that will be used.
This includes our own dance data, reorganised by `GUV_organisedata.R` as `colour_dance_reorg.csv`, as well as data from Edrich *et al.* (1979) on the relationship between accuracy (mean vector length), wavelength and intensity.
This was manually extracted from Fig.
2 and saved in `"Edrich1979_brightness-accuracy.csv`.

```{r Load data}
#Dance angle data
cd = read.table(file = '1Data/colour_dance_reorg.csv', 
                header = T, 
                sep  = ',')
#Accuracy data
ed = read.table(file = '1Data/Edrich1979_brightness-accuracy.csv', 
                header = T, 
                sep  = ',')
#Inspect each dataset
print(cd)
print(ed)
```

## Format data

Ensure that each variable has the correct data format for subsequent analysis.

```{r Data formatting}
cd = within(cd,
            {
              ID = as.factor(ID) # beedance identifier as a factor
              date = as.factor(date) # date as a factor
              signed_angle = Mod360.180(bearing)  # bearing between -180 and 180
              angle = circular(rad(signed_angle),# bearing between -pi and pi
                               rotation = 'clock') # circular format suppresses later warnings
            }
)
u_id = with(cd, unique(ID)) # unique beedances
length(u_id)#169 beedances
```

Identify dances in which all four stimulus conditions were used.

```{r full condition}
luc = sapply(u_id,
              FUN = IndCond,
              dt = cd)

#Most individuals that made it to bright green 
full_ids = u_id[luc == 4]
length(full_ids)#19 individuals
#extract just those individuals

#find data for full condition individuals
full_cd = subset(cd, ID %in% full_ids)
```

## Plot the raw dance data

Plot all 169 dances

```{r Raw dances, echo=FALSE, warning=FALSE}
#TODO make function
#set up a sequence for the axes
xc = seq(from = -pi, to = pi-1e-16, length.out = 1e3)
par(mfrow = c(2, 2),
    mar = c(0,0,0,0))
par(pty = 's')

#add legend to 1st page
plot(x = NULL,
     xlim = c(-1,1),
     ylim = c(-1,1),
     pch = 19,
     axes = FALSE,
     xlab = '',
     ylab = '',
     main = ''
)
legend(x = 'center',
       legend = c('Green Bright',
                  'Green Dim',
                  'UV Bright',
                  'UV Dim',
                  'Sun azimuth (North = Up)'),
       col = c('green',
               'darkgreen',
               'magenta',
               'purple',
               'orange'),
       pch = c(21,21,21,21,NA),
       lty = c(NA, NA, NA, NA, 1),
       lwd = c(2,2,2,2,3)
       )
#loop through individuals
for(ii in u_id)
{
  mnv_gh = Mvec(subset(cd, ID %in% ii & colour %in% 'g' & brightn %in% 'h')$angle)
  mnv_gl = Mvec(subset(cd, ID %in% ii & colour %in% 'g' & brightn %in% 'l')$angle)
  mnv_uh = Mvec(subset(cd, ID %in% ii & colour %in% 'u' & brightn %in% 'h')$angle)
  mnv_ul = Mvec(subset(cd, ID %in% ii & colour %in% 'u' & brightn %in% 'l')$angle)
  plot(x = NULL,
       xlim = 100*c(-1,1),
       ylim = 100*c(-1,1),
       pch = 19,
       axes = FALSE,
       xlab = '',
       ylab = '',
       main = ''
       )
  abline(a = 0, b = 1, col = 'gray90')
  abline(a = 0, b = -1, col = 'gray90')
  abline(h = 0, v = 0, col = 'gray75')
  lines(x = 20*sin(xc), y = 20*cos(xc), lty = 3)
  lines(x = 40*sin(xc), y = 40*cos(xc), lty = 3)
  lines(x = 60*sin(xc), y = 60*cos(xc), lty = 3)
  lines(x = 80*sin(xc), y = 80*cos(xc), lty = 3)
  text(x = 1:4*20,
       y = c(0,0,0,0),
       labels = paste(' run', 1:4*20), 
       cex = 0.3,
       adj = c(0,1))
  with(subset(cd, ID %in% ii),
       {
  lines(x = c(0, 100*sin(sun_az_rad)),
        y = c(0, 100*cos(sun_az_rad)),
        col = adjustcolor('orange',alpha.f = 0.5),
        lwd = 2)
  points(x = run*sin(as.numeric(angle)),
         y = run*cos(as.numeric(angle)),
         bg = gray(level = 1.0,
                   alpha =  0.4),
         col = c('green', 'darkgreen', 'magenta', 'purple')
                [ifelse(colour %in% 'g',
                        yes = ifelse(brightn %in% 'h', yes = 1, no = 2),
                        no = ifelse(brightn %in% 'h', yes = 3, no = 4))],
        pch = 21,
        lwd = 2
       )
       }
  )
  lines(x = c(0,100*sin(mnv_gh['mu'])*mnv_gh['rho']), 
        y = c(0,100*cos(mnv_gh['mu'])*mnv_gh['rho']), 
        col = 'green', 
        lwd = 1)
  lines(x = c(0,100*sin(mnv_gl['mu'])*mnv_gl['rho']), 
        y = c(0,100*cos(mnv_gl['mu'])*mnv_gl['rho']), 
        col = 'darkgreen', 
        lwd = 1)
  lines(x = c(0,100*sin(mnv_uh['mu'])*mnv_uh['rho']), 
        y = c(0,100*cos(mnv_uh['mu'])*mnv_uh['rho']), 
        col = 'magenta', 
        lwd = 1)
  lines(x = c(0,100*sin(mnv_ul['mu'])*mnv_ul['rho']), 
        y = c(0,100*cos(mnv_ul['mu'])*mnv_ul['rho']), 
        col = 'purple', 
        lwd = 1)
  mtext(ii,side = 3,line = -2)
}
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Accuracy effects

Calculate the mean vectors for each set of waggle-runs under each lighting condition.

```{r Summarise, echo=FALSE}
#calculate mean vectors
mean_vectors = aggregate(angle~ID*brightn*colour+sun_az, # N.B. Including sun azimuth drops some cases without a time stamp
                         data = cd,
                         FUN = rho.circular
)
#correct names
mean_vectors = within(mean_vectors,
                      {mean_vector = angle; rm(angle)} # anlge now indicates a mean vector, not an angle
)
#add kappa
mle_estimates = aggregate(angle~ID*brightn*colour+sun_az,
                          data = cd,
                          FUN = MLE_est
)
#add to the summary table and 
#calculate inverse softplus kappa
mean_vectors = within(mean_vectors,
                      {
                        mu = deg(mle_estimates$angle[,'mu'])
                        kappa = mle_estimates$angle[,'kappa']
                        iskappa  = inv_softplus(kappa)
                      }
)
#find the full condition individuals
mean_vectors_full = subset(mean_vectors,
                           subset = ID %in% full_ids)
```

### Effect of brightness on mean vector length

#### Entire dataset

The first thing we notice is that in the low intensity conditions the accuracy is generally lower than for the high intensity conditions.

```{r Plot mean vectors, echo=FALSE}
boxplot(mean_vector~colour*brightn,
        data = mean_vectors,
        ylim = c(0,1),
        col = adjustcolor(c('green2',
                            'purple',
                            'darkgreen',
                            'purple4'), alpha.f = 0.5),
        outline = FALSE)
stripchart(mean_vector~colour*brightn,
           data = mean_vectors,
           add = TRUE,
           vertical = TRUE,
           method = 'jitter',
           col = adjustcolor(c('green2',
                               'purple',
                               'darkgreen',
                               'purple4'), alpha.f = 0.5),
           bg = gray(level = 0,
                     alpha = 0.1),
           pch = 21)
#Identify stimuli
stim = apply(X = expand.grid(c = c('g','u'),b = c('h','l')),
             FUN = paste,
             collapse = '',
             MARGIN = 1)
#Plot lines linking dances by the same individual
invisible(
  {
    lapply(X = u_id,
           FUN = Plt_mvec
           )
  })
abline(h = c(0,1))
```

#### Within individuals (balanced cases only)

This effect is even clearer if we look at just the full condition dances.

```{r Plot full condition mean vectors}
stripchart(mean_vector~colour*brightn,
           data = mean_vectors_full,
           # add = TRUE,
           vertical = TRUE,
           method = 'jitter',
           jitter = 0.02,
           ylim = c(0,1),
           col = adjustcolor(c('green2',
                               'purple',
                               'darkgreen',
                               'purple4'), alpha.f = 0.5),
           bg = gray(level = 0,
                     alpha = 0.1),
           pch = 21)
#Plot lines linking dances by the same individual
invisible(
  {
    lapply(X = full_ids,
           FUN = Plt_mvec
           )
  })
abline(h = c(0,1))
```

### Comparison with Edrich *et al.* (1979)

How does this compare with previous experiments in which a light spot (x°) was presented through bandpass filters and ND filters to produce different intensities?
In that data, there was a clear sigmoidal relationship between intensity and accuracy (mean vector length).

#### Plot raw data from Edrich *et al.* (1979) Fig. 2

```{r Edrich raw data, echo=FALSE}
with(ed,
     {
     plot(x = log10(intensity),
          y = accuracy,
          xlim = c(9,14),
          ylim = c(0,1),
          main = 'Data from Edrich et al. 1979, Fig. 2',
            xlab = 'log10(intensity) (photons/cm2/s)',
            ylab = 'accuracy (r)',
          pch = 21,
          cex = 1.5,
          col = 'black',
          bg = sapply(paste(wavelength),
                       FUN = switch,
                       `354` = 'purple',
                       `429` = 'blue',
                       `535` = 'darkgreen',
                       'red')
          )
     abline(h = c(0,1))
     }
     )

legend(x = 'bottomright',
       legend = paste(sort(unique(ed$wavelength),
                     decreasing = TRUE), 'nm'),
       col = c('darkgreen','blue', 'purple'),
       pch = c(20, 20, 20))
```

This is best modelled with a nonlinear (psychometric model) following the beta distribution (for variables bounded between zero and one).

#### Reorganise data

```{r Prepare data for model fitting}
ed = within(ed,
            {
              logit_accuracy = qlogis(accuracy)
              log10_intensity = log10(intensity)
              colour = sapply(paste(wavelength),
                              FUN = switch,
                              `354` = 'UV',
                              `429` = 'blue',
                              `535` = '0green',
                              NA)
            }
            )
```

#### Fit psychometric curves

For each light stimulus, fit a sigmoid curve bounded between baseline accuracy (shared across stimuli) and maximum accuracy (`1 - lapses`).
Inflection point and 80%-rise-region width vary between stimuli, maximising posterior probability for a beta distribution in `(0,1)`.
To do this, we first define a nonlinear formula in `brms`.

```{r Set up formula}

#set up model fit
formula_beta = bf(
  formula = accuracy ~ 
    inv_logit(LogitBase) + (1 - inv_logit(LogitLapse) - inv_logit(LogitBase) ) *#curve region
    inv_logit( 4.39*(log10_intensity - Inflex) / exp(LogWidth) ) , #inflection-width curve
  # for each of these parameters, we can set up a separate formula 
  # that describes how to predict them from the data 
  #Base rate of correct choices: "
  LogitBase ~ 1, #Base rate of correct choices: "~ 1" gives the instruction "estimate the mean across all data"
  #Lapse rate on a log(odds) scale:
  LogitLapse ~ colour, #this is similar to the formula in our LMM example
  #inflection point of the initial curve:
  Inflex ~ colour, #N.B. this is similar to the intercept, so it does not include effects of stimulus level
  #log 80% width of the curve:
  LogWidth ~ colour, #N.B. this is similar to the slope, so all of its effects depend on stimulus level
  family = Beta(link = "identity", link_phi = 'log'),
  nl = TRUE)#the joint distribution for these parameters is undefined, and therefore the parameters themselves are "nonlinear"

```

Since this is a nonlinear model (there are terms both inside and outside the logistic transform) it is important to establish prior distributions for all parameters.
These ensure that parameter estimates explore credible ranges during fitting.
The baseline is expected to be around $r = 0.27$, while the maximum accuracy should be around $r = 0.95$, and their estimates should not converge (that would flatten the curve, or reverse its direction).
The inflection points are expected around $10^{11}$ photons cm$^{-2}$ s$^{-1}$, with a rise-region width of $10^{3}$ photons cm$^{-2}$ s$^{-1}$.
All colour effects may vary by a few orders of magnitude.

```{r}

#set up priors
prior_beta = get_prior(formula = formula_beta,
                     data = ed)
#set the prior distribution
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogitBase' &
                      coef %in% 'Intercept' 
                  ] = 'normal(-1,1)' #a normal distribution centred on plogis(-1) = 0.27
                  })
#this prior is automatically bounded between 0 and 1, 
#Lapse rate priors ---------------------------------------------------
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogitLapse' &
                      coef %in% 'Intercept' 
                  ] = 'normal(-3,2)' #a normal distribution centred on -3
                  })

#for all other fixed effects on lapse rate, we'll suggest values around 0 (no effect)
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogitLapse' &
                      coef %in% 'colourblue' 
                  ] = 'normal(0,2)' #a normal distribution:mean 0, sd 3
                  })
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogitLapse' &
                      coef %in% 'colourUV' 
                  ] = 'normal(0,2)' #a normal distribution:mean 0, sd 3
                  })

#Inflection point priors ---------------------------------------------
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'Inflex' &
                      coef %in% 'Intercept'
                  ] = 'normal(11,2)' #a normal distribution:mean 11, sd 2
                  })
#for all coefficients, we'll suggest values around 0 (no effect)
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'Inflex' &
                      coef %in% 'colourblue' 
                  ] = 'normal(0,2)' #a normal distribution:mean 0, sd 2
                  })

prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'Inflex' &
                      coef %in% 'colourUV' 
                  ] = 'normal(0,2)' #a normal distribution:mean 0, sd 2
                  })

#Rise region width priors --------------------------------------------
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogWidth' &
                      coef %in% 'Intercept' 
                  ] = 'normal(1,2)' #a normal distribution:mean 1, sd 3
                  })
#for all coefficients, we'll suggest values around 0 (no effect)
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogWidth' &
                      coef %in% 'colourblue'
                  ] = 'normal(0,2)' #a normal distribution:mean 0, sd 3
                  })
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogWidth' &
                      coef %in% 'colourUV'
                  ] = 'normal(0,2)' #a normal distribution:mean 0, sd 3
                  })
```

Fit a psychometric model to the data using the nonlinear formula and specified prior distributions.
`4000` iterations on `4` chains takes less than one minute.

```{r Fit model}
#takes less than 1 minute
system.time(
  {
    beta_fit = brm( formula = formula_beta, # using our nonlinear formula
                     data = ed, # our data
                     prior = prior_beta, # our priors 
                     iter = 4000, # long run for 4000 iterations
                     chains = 4, # 4 chains in parallel
                     cores = 4, # on 4 CPUs
                     refresh = 0, # don't echo chain progress
                     control = list(adapt_delta = 0.95), #finer sampling
                     backend = 'cmdstanr') # use cmdstanr (other compilers broken)
  }
)
```

#### Plot and compare data from Edrich and Kühn.

Summarise the data from full cases in the Kühn data using quantiles (minimum, 25%, median, 75% and maximum).

```{r Collect relevant data from Kühn data}
#Examine quantiles
meanQ = aggregate(mean_vector ~ colour*brightn, 
                  data = mean_vectors_full,
                  FUN = quantile)
#Extract different quantiles as separate columns in a data frame
meanQ = data.frame(cbind(meanQ[1:2], meanQ$mean_vector) )
#Rename
names(meanQ) = c('colour', 'brightn', paste0('q',0:4))
#Subset by condition
gh_meanQ = subset(meanQ, colour %in% 'g' & brightn %in% 'h')
gl_meanQ = subset(meanQ, colour %in% 'g' & brightn %in% 'l')
uh_meanQ = subset(meanQ, colour %in% 'u' & brightn %in% 'h')
ul_meanQ = subset(meanQ, colour %in% 'u' & brightn %in% 'l')

```

Extract the model predictions by summarising parameter estimates from the sampling phases of all chains.
To identify specific combinations of parameter values, use the `conditional_effects` function, which organises all simultaneous estimates into a data frame.

```{r Extract model predictions}

# #N.B. for plotting we will need to also define "inv_logit" in our the R environment
inv_logit = inv_logit_scaled # this can be found in brms, but with an additional specifier

#summary of parameter estimates
full_sm_beta = summary(beta_fit,
                  robust = TRUE)#use the median estimate

#Extract fixed effects estimates
full_fix_beta = full_sm_beta$fixed
#extract rownames
full_fix_rn_beta = rownames(full_fix_beta)

# stimulus levels to add to the dataset to be able to measure this effect.
full_estimates_beta = full_fix_beta$Estimate
names(full_estimates_beta) = full_fix_rn_beta

# to predict all effects, we can use the 'posterior_epred' method
#by default 100 predictions per continuous variable (but fewer for their interactions)
#This takes <2 seconds
system.time(
  {
    full_cond_beta =brms::conditional_effects(beta_fit, 
                                         method = 'posterior_epred', # posterior epred not working
                                         cores =  parallel::detectCores()-1,
                                         effects = c('log10_intensity:colour'),
                                         plot = FALSE
    )
  }
)
#extract predictions
pred_data_beta = full_cond_beta$`log10_intensity:colour`

#Green light curve
g_curve = 
with(data.frame(t(full_estimates_beta)),
     {
       list(Base = plogis(LogitBase_Intercept),
            Maxima = 1- plogis(LogitLapse_Intercept), 
            Inflection = Inflex_Intercept,  
            Width = exp(LogWidth_Intercept)  
       )
     }
)

#UV light curve
u_curve = 
with(data.frame(t(full_estimates_beta)),
     {
       list(Base = plogis(LogitBase_Intercept),
            Maxima = 1- plogis(LogitLapse_Intercept + LogitLapse_colourUV), 
            Inflection = Inflex_Intercept + Inflex_colourUV,  
            Width = exp(LogWidth_Intercept + LogWidth_colourUV)  
       )
     }
)



```

Plot the raw data alongside the predictions for the data from Edrich *et al*., (1979) and a boxplot summary of the data data from Kühn (2017).

```{r Plot model predictions}

#Prepare a sequence for plotting
xxl = seq(from = 9, to = 16,
          length.out = 1e3)

#plot only the stimulus types also used by Kühn.
ed_ug = subset(ed, wavelength != 429)

#plot each stimulus type
with(ed_ug,
     {
       plot(x = log10_intensity,
            y = accuracy,
            main = 
'Comparison of data from Edrich et al. 1979
with Kühn, 2017',
            xlab = 'log10(intensity) (photons/cm2/s)',
            ylab = 'accuracy (r)',
            ylim = c(0,1),
            xlim = c(9,16),
            pch = 21,
            cex = 1.5,
            col = 'black',
            bg = sapply(paste(wavelength),
                        FUN = switch,
                        `354` = 'purple',
                        `429` = 'blue',
                        `535` = 'darkgreen',
                        'red')
       )
       abline(h = c(0,1))
       abline(v = log10(4*c(1e12,1e14)) )
     }
)
#Add the estimated range of the curve for the green stimulus
with(data.frame(t(full_estimates_beta)),
     {
  abline(h = c(0,1,
               plogis(LogitBase_Intercept),
               1- plogis(LogitLapse_Intercept) ), 
         lty = c(1,1,3,3)
         )
     }
)

#plot total 95% prediction intervals
with(subset(pred_data_beta, colour == '0green'), 
     {
       polygon(x = c(sort(log10_intensity), rev(sort(log10_intensity))), 
               y = c(lower__[order(log10_intensity)],
                     rev(upper__[order(log10_intensity)])
               ), 
               col = adjustcolor('green', alpha.f = 25/256),
               border = NA,
               lwd = 0.1
       )
     }
)
with(subset(pred_data_beta, colour == 'UV'), 
     {
       polygon(x = c(sort(log10_intensity), rev(sort(log10_intensity))), 
               y = c(lower__[order(log10_intensity)],
                     rev(upper__[order(log10_intensity)])
               ), 
               col = adjustcolor('magenta', alpha.f = 25/256),
               border = NA,
               lwd = 0.1
       )
     }
)

#plot the median prediction lines
with(g_curve, 
     lines(x = xxl,
           y = Base + (Maxima  - Base) * 
             plogis(4.39*(xxl - Inflection)/Width), 
           col = 'darkgreen',
           lwd = 3,
           lty = 3)
)
with(u_curve, 
     lines(x = xxl,
           y = Base + (Maxima  - Base) * 
             plogis(4.39*(xxl - Inflection)/Width), 
           col = 'purple4',
           lwd = 3,
           lty = 3)
)

#plot observed data from Kühn as boxplots showing the quartiles, median and total range.
#Green High intensity
QBox(quant = subset(meanQ, colour %in% 'g' & brightn %in% 'h'),
     x = log10(4e14),
     offset = -0.1,
     cols = c('green', 'darkgreen')
    )

#Green Low intensity
QBox(quant = subset(meanQ, colour %in% 'g' & brightn %in% 'l'),
     x = log10(4e12),
     offset = -0.1,
     cols = c('green', 'darkgreen')
    )

#UV High intensity
QBox(quant = subset(meanQ, colour %in% 'u' & brightn %in% 'h'),
     x = log10(4e14),
     offset = 0.1,
     cols = c('magenta', 'purple')
    )

#UV Low intensity
QBox(quant = subset(meanQ, colour %in% 'u' & brightn %in% 'l'),
     x = log10(4e12),
     offset = 0.1,
     cols = c('magenta', 'purple')
    )

legend(x = 'bottomright',
       legend = c(
                   paste('Edrich', 
                        sort(unique(ed_ug$wavelength),
                           decreasing = TRUE), 'nm'),
                   paste('  Kühn', c(528, 365), 'nm' ) 
       ),
       col = c('darkgreen', 'purple', 'green', 'magenta'),
       pch = c(20, 20, 15, 15),
       lty = c(3,3,NA, NA),
       lwd = 3,
       cex = 0.7)
```

It would be difficult to explain the orientation accuracy for the high intensity ($4*10^{14}$) UV stimulus from the accuracy curve for Edrich's UV data, but it would fit well with the maximum accuracy observed for green light.
In Edrich's study, the green light was interpreted as the sun (dances at correct angle between sun and feeder), whereas the UV light was interpreted as a patch of sky near the anti-sun.
In Kühn's data the stimuli were smaller (assuming a viewing distance of ≈200mm, an SMD with diameter 3.5mm would appear $deg(2*atan2(y = 3.5/2,x = 200)) =$ 1.0° across) than Edrich's (2.3°), which may have resulted in them being perceived as dimmer (left shift in this graph).
Edrich $et~al.$ (1979) mention that the particular relationship with light intensity was variable.
They also mention that the relative relationships remained constant, so we may also observe the same relationship but right shifted.

> Unfortunately, the scatter of the dance directions as a function of light intensity proved to be unstable over long periods of time; the absolute values changed from day to day.
> However, these changes were such that curves for adjacent wavelengths were shifted in parallel - the relative sensitivities remained constant.
> --- Edrich et al. (1979), p154

# Sun compass orientation

Since most dances were at least somewhat oriented ($r>0.5$) we can assume that the mean angles of these dances typically align with an intended dance direction.
If most dances are for the same geographic location, we should observe dances as the same fixed angle to the stimulus, interpreted either as the sun or the anti-sun.

```{r Calculate and plot angles relative expected sun azimuth}

sun_diff_gh = sapply(X = u_id,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'g',
                    br = 'h')
sun_diff_gl = sapply(X = u_id,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'g',
                    br = 'l')
sun_diff_uh = sapply(X = u_id,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'u',
                    br = 'h')
sun_diff_ul = sapply(X = u_id,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'u',
                    br = 'l')


par(mfrow = c(2,2), mar = c(0,0,0,0))
PCfun(angles = unlist(sun_diff_gh),
      col = 'green',
      shrink = 2.0,
      side = 3,
      title = 'Green Bright')
PCfun(angles = unlist(sun_diff_gl),
      col = 'darkgreen',
      shrink = 2.0,
      side = 3,
      title = 'Green Dim')
PCfun(angles = unlist(sun_diff_uh),
      col = 'magenta',
      shrink = 2.0,
      side = 3,
      title = 'UV Bright')
PCfun(angles = unlist(sun_diff_ul),
      col = 'purple',
      shrink = 2.0,
      side = 3,
      title = '           UV Dim')
```

For the bright stimuli the dances generally indicate South, but with a very wide range.
(or possibly clusters to the South-West and South-East).
For the dim stimuli there is less of a clear pattern, particularly for UV dim.

We can also restrict this analysis to just the dances with all conditions.

```{r Angles relative expected sun azimuth for full conditions}

full_sun_diff_gh = sapply(X = full_ids,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'g',
                    br = 'h')
full_sun_diff_gl = sapply(X = full_ids,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'g',
                    br = 'l')
full_sun_diff_uh = sapply(X = full_ids,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'u',
                    br = 'h')
full_sun_diff_ul = sapply(X = full_ids,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'u',
                    br = 'l')


par(mfrow = c(2,2), mar = c(0,0,0,0))
PCfun(angles = unlist(full_sun_diff_gh),
      col = 'green',
      shrink = 1.5,
      side = 3,
      title = 'Green Bright')
PCfun(angles = unlist(full_sun_diff_gl),
      col = 'darkgreen',
      shrink = 1.5,
      side = 3,
      title = 'Green Dim')
PCfun(angles = unlist(full_sun_diff_uh),
      col = 'magenta',
      shrink = 1.5,
      side = 3,
      title = 'UV Bright')
PCfun(angles = unlist(full_sun_diff_ul),
      col = 'purple',
      shrink = 1.5,
      side = 3,
      title = 'UV Dim')
      
```

When restricting to only the balanced cases, the pattern becomes clearer.
For the most oriented dances to the bright UV stimulus there are two clusters to the South-West and South-East.
These are likely also present for the bright green stimulus.
The dim green stimulus appears to align somewhat with the bright stimuli.
The dim UV stimulus shows no apparent patten, though a slight southwards bias.

# Inspect dance angles relative to the stimulus

## Mean dance angles relative to the stimulus

Since dances do not point to a specific geographic location, it is possible that they advertise a variety of directions relative to the sun, or are not menotactic but phototactic or uniformly oriented.
We can identify phototactic orientation from the angle relative to the stimulus.

```{r Plot dance angle relative to stimulus}
par(mfrow = c(2,2), mar = c(0,0,0,0))
PCfun(angles = subset(mean_vectors_full, 
                      subset = colour %in% 'g' & brightn %in% 'h')$mu,
      col = 'green',
      shrink = 1.5,
      title = 'Green Bright')
PCfun(angles = subset(mean_vectors_full, 
                      subset = colour %in% 'g' & brightn %in% 'l')$mu,
      col = 'darkgreen',
      shrink = 1.5,
      title = 'Green Dim')
PCfun(angles = subset(mean_vectors_full, 
                      subset = colour %in% 'u' & brightn %in% 'h')$mu,
      col = 'magenta',
      shrink = 1.5,
      title = 'UV Bright')
PCfun(angles = subset(mean_vectors_full, 
                      subset = colour %in% 'u' & brightn %in% 'l')$mu,
      col = 'purple',
      shrink = 1.5,
      title = 'UV Dim')
```

While there are some dances towards or near to the stimulus for most conditions, there is a wide spread, suggesting a factor other than phototaxis sets dance directions.
N.B.
for the dim stimuli where the widest spread is observed, the dances are also less accurate, which may increase variation in mean dance angle.

## Change in mean dance angle between stimuli

### Change arcs

To determine whether these dances interpret the stimuli as the sun, anti-sun or something else, we would need to check the differences in dance angle between the conditions.
It is also worth considering any differences in accuracy between dances to each condition, since this would influence the variation in angular change.
These can be visualised as arcs.

```{r Plot change arcs}
#TODO make a function for this
par(mfrow = c(2,2), mar = c(0,0,0,0))
OpenCplot(n_sample = with(full_cd, median(run)/4) )
mtext(text = 'Green Bright to Green Dim',
      side = 1,
      line = -1,
      cex = 0.5)
#open the loop
for(ii in mean_vectors_full$ID)
{
with(subset(mean_vectors_full,
            subset = ID %in% ii),
     {
      DiffArc(a1 = mu[colour %in% 'g' & 
                      brightn %in% 'h'],
              a2 = mu[colour %in% 'g' & 
                        brightn %in% 'l'],
              r1 = mean_vector[colour %in% 'g' & 
                              brightn %in% 'h'],
              r2 = mean_vector[colour %in% 'g' & 
                                brightn %in% 'l'],
              col1 = 'green',
              col2 = 'darkgreen'
              )
     }
    )
}

OpenCplot(n_sample = with(full_cd, median(run)/4) )
mtext(text = 'UV Bright to UV Dim',
      side = 1,
      line = -1,
      cex = 0.5)
for(ii in mean_vectors_full$ID)
{
with(subset(mean_vectors_full,
            subset = ID %in% ii),
     {
      DiffArc(a1 = mu[colour %in% 'u' & 
                      brightn %in% 'h'],
              a2 = mu[colour %in% 'u' & 
                        brightn %in% 'l'],
              r1 = mean_vector[colour %in% 'u' & 
                              brightn %in% 'h'],
              r2 = mean_vector[colour %in% 'u' & 
                                brightn %in% 'l'],
              col1 = 'magenta',
              col2 = 'purple'
              )
     }
    )
}
OpenCplot(n_sample = with(full_cd, median(run)/4) )
mtext(text = 'Green Bright to UV Bright',
      side = 1,
      line = -1,
      cex = 0.5)
for(ii in mean_vectors_full$ID)
{
with(subset(mean_vectors_full,
            subset = ID %in% ii),
     {
      DiffArc(a1 = mu[colour %in% 'g' & 
                      brightn %in% 'h'],
              a2 = mu[colour %in% 'u' & 
                        brightn %in% 'h'],
              r1 = mean_vector[colour %in% 'g' & 
                              brightn %in% 'h'],
              r2 = mean_vector[colour %in% 'u' & 
                                brightn %in% 'h'],
              col1 = 'green',
              col2 = 'magenta'
              )
     }
    )
}
OpenCplot(n_sample = with(full_cd, median(run)/4) )
for(ii in mean_vectors_full$ID)
{
with(subset(mean_vectors_full,
            subset = ID %in% ii),
     {
      DiffArc(a1 = mu[colour %in% 'g' & 
                      brightn %in% 'l'],
              a2 = mu[colour %in% 'u' & 
                        brightn %in% 'l'],
              r1 = mean_vector[colour %in% 'g' & 
                              brightn %in% 'l'],
              r2 = mean_vector[colour %in% 'u' & 
                                brightn %in% 'l'],
              col1 = 'darkgreen',
              col2 = 'purple'
              )
     }
    )
}
mtext(text = 'Green Dim to UV Dim',
      side = 1,
      line = -1,
      cex = 0.5)
```

### Differences in mean angle

```{r Plot changes in mean dance angle}
mu_diff_gl = sapply(X = full_ids,
                    FUN = MuDiff,
                    dt = mean_vectors_full,
                    cl = 'g',
                    br = 'l')
mu_diff_uh = sapply(X = full_ids,
                    FUN = MuDiff,
                    dt = mean_vectors_full,
                    cl = 'u',
                    br = 'h')
mu_diff_ul = sapply(X = full_ids,
                    FUN = MuDiff,
                    dt = mean_vectors_full,
                    cl = 'u',
                    br = 'l')


#Other contrasts
mu_diff_uhl = sapply(X = full_ids,
                    FUN = MuDiff,
                    dt = mean_vectors_full,
                    cl = 'u',
                    br = 'l',
                    ref_cl = 'u',
                    ref_br = 'h')
mu_diff_gul = sapply(X = full_ids,
                    FUN = MuDiff,
                    dt = mean_vectors_full,
                    cl = 'u',
                    br = 'l',
                    ref_cl = 'g',
                    ref_br = 'h')

par(mfrow = c(2,2), mar = c(0,0,0,0))
PCfun(angles = unlist(mu_diff_gl),
      col = 'darkgreen',
      shrink = 1.5,
      title = 'Green Dim - Green Bright')
PCfun(angles = unlist(mu_diff_uhl),
      col = 'purple',
      shrink = 1.5,
      title = 'UV Dim - UV Bright')
PCfun(angles = unlist(mu_diff_uh),
      col = 'gray40',
      shrink = 1.5,
      title = 'UV Bright - Green Bright')
PCfun(angles = unlist(mu_diff_gul),
      col = 'gray25',
      shrink = 1.5,
      title = 'UV Dim - Green Dim')

```

# Estimate turn effects

## Find fit maximum likelihood turn distributions

For each change distribution, find the maximum likelihood von Mises distribution.
For the UV high to UV low change, this will be a bimodal distribution.

```{r Organise heading changes}
#collect all heading differences
pair_diffs_lst = lapply(X = list(mu_diff_gl,
                                 mu_diff_uh,
                                 mu_diff_ul,
                                 mu_diff_gul), 
                        FUN = unlist) 
#Give a simpler name
names(pair_diffs_lst) = c('green_hl',
                          'uvg_h',
                          'uvg_hl',
                          'uvg_l')
#Provide appropriate data format
angle_unit = 'degrees' 
angle_rot = 'clock'
#make sure they are in circular format
pair_diffs_lst = lapply(X = pair_diffs_lst,
                        FUN = circular,
                        units = angle_unit,
                        rotation = angle_rot)
```

Fit the maximum likelihood distribution for no difference between pairs, differences between pairs, and differences between pairs with two means.
If angles shift in one direction between trials, the "difference" distribution should have significantly higher likelihood.
If angles shift in two directions between pairs, then the "two means" distribution should be the best fit.

```{r Fit maximum likelihood von Mises distributions}
#TODO neaten up!
#Distribution for same means
ml_same_lst = lapply(X = pair_diffs_lst,
                     mle.vonmises,
                     bias = TRUE,
                     mu = circular(x = 0,
                                   units = angle_unit,
                                   rotation = angle_rot)
)
#Distribution for different means
ml_diff_lst = lapply(X = pair_diffs_lst,
                     mle.vonmises,
                     bias = TRUE)
#Distribution for multiple different means
mod_diff_lst = lapply(X = pair_diffs_lst,
                      FUN = circ_mle,#converts circular format data to radians
                      niter = 1e4)#some stochastic variability for green stim
#Forced multiple different means
bim_diff_lst = lapply(X = pair_diffs_lst,
                      FUN = circ_mle,#converts circular format data to radians
                      exclude = c('M1','M2A', 'M2B', 'M2C'),#no uniform or unimodal
                      niter = 1e4)#some stochastic variability for green stim


#Extract the parameters of the model with the lowest AIC
mod_diff_par = t( sapply(X = mod_diff_lst,
                      FUN = MD_extract) )
bim_diff_par = t( sapply(X = bim_diff_lst,
                      FUN = MD_extract) )

# Calculate likelihood 
#calculate all in list
all_ll = mapply(FUN = LLcalc,
                ml = c(ml_same_lst, ml_diff_lst), #apply to both sets of ML estimates
                angles = pair_diffs_lst, #repeat datasets
                au = angle_unit, #important, use the same frame of reference as the original
                ar = angle_rot)

#inspect resulting parameters
mod_diff_results = do.call(rbind, c(ml_same_lst, ml_diff_lst))
#add easy to parse names
rownames(mod_diff_results) = paste( sort( #distinguish same and different mean
                          rep(x = c( 'same', 'diff'), 
                              times =  length(pair_diffs_lst) ),
                          decreasing = TRUE
                          ),
                       rep(x = names(pair_diffs_lst),
                           times = 2) 
                       )
#display version
rspar = mod_diff_results[,c('mu', 'kappa', 'se.mu', 'se.kappa')]
rspar = cbind(rspar, loglikelihood = all_ll)
rspar = apply(X = rspar, MARGIN = 2, FUN = unlist)
#reorder for easy comparison
rspar = rspar[c(rbind(1:4, 1:4+4)),]#put the different directly after the same
rspar[,3:4] = NA
rspar = cbind(rspar, weight1 = 1.0)
colnames(rspar) = colnames(bim_diff_par)[c(1:4, 6:5)]
all_res = rbind(data.frame(rspar),
                data.frame(bim_diff_par) )
all_res = apply(all_res, MARGIN = 2, FUN = unlist)
rownames(all_res[9:12,]) = paste('multi', rownames(bim_diff_par))
all_res = all_res[c(1:2, 9, 3:4, 10, 5:6, 11, 7:8, 12),]

all_res = within(data = data.frame(all_res),
                 {
                   model = c('same',
                             'diff',
                             'multi')
                   dataset = c(sapply(X = names(pair_diffs_lst),
                                    FUN = rep,
                                    times = 3) )
                 }
                )
# print(rspar, digits = 3)
# print(mod_diff_par, digits = 3)
# print(bim_diff_par, digits = 3)
print(all_res, digits = 3)

```

As expected, the bimodal models fitted by `circ_mle` have the highest likelihood, followed by the models for differences and then the models for no difference.
The difference models require at ≥1 more parameter, so should increase likelihood by \>1 order of magnitude to be appropriate.
The bimodal model require ≥2 more parameters, so should increase likelihood by \>2 log units.

## Estimate confidence intervals

For each mean, simulate a distribution of mean estimates

```{r Calculate the confidence intervals}
#estimate CI for unimodal von Mises models
mle_CI = with(data.frame(mod_diff_results), 
              {
              mapply(FUN = CI_vM, 
                     angles = pair_diffs_lst, 
                     m1 = mu, 
                     k1 = kappa, 
                     w1 = 1,
                     alternative = 'two.sided',#not a hypothesis test (symmetrical)
                     n = 1e4)#takes ≈30s for 10 000 simulations
                }
              )
#reorder for comparison like the original models
mle_CI =  t(mle_CI)[c(rbind(1:4, 1:4+4)),]
rownames(mle_CI) = rownames(rspar)
colnames(mle_CI) = c('lower','median','upper')

#estimate CI for unimodal and bimodal von Mises models
mod_CI = with(data.frame(mod_diff_par), 
              {
              mapply(FUN = CI_vM, 
                     angles = pair_diffs_lst, 
                     m1 = mu1, 
                     k1 = kappa1, 
                     m2 = mu2, 
                     k2 = kappa2, 
                     w1 = weight1,
                     alternative = 'two.sided',#not a hypothesis test (symmetrical)
                     # SIMPLIFY = FALSE,
                     n = 1e4)#takes ≈15 for 10 000 simulations
                }
              )
#estimate CI for unimodal and bimodal von Mises models
bim_CI = with(data.frame(bim_diff_par), 
              {
              mapply(FUN = CI_vM, 
                     angles = pair_diffs_lst, 
                     m1 = mu1, 
                     k1 = kappa1, 
                     m2 = mu2, 
                     k2 = kappa2, 
                     w1 = weight1,
                     SIMPLIFY = FALSE,
                     alternative = 'two.sided',#not a hypothesis test (symmetrical)
                     n = 1e4)#takes ≈15 for 10 000 simulations
                }
              )
#collapse to printable table
mod_CI_tb = Table_vCI(mod_CI)
bim_CI_tb = Table_vCI(bim_CI)
print(mle_CI, digits = 3)
print(mod_CI_tb, digits = 3)

```

We see here that while the confidence interval for the difference distribution between the two green conditions and the two bright conditions nearly overlap with 0° (and the 0° distribution CI overlap with the mean difference), for differences between green lights and dim UV there is a bimodal distribution with CI that to not overlap with 0°.
N.B.
For *one-sided* calculations, the CI do overlap with 0° for green conditions, bright conditions and the leftmost mean for dim conditions.
### Plot difference CI

```{r Plot CI}
#TODO make bimodal version of PCfun
par(mfrow = c(2,2), mar = c(0,0,0,0))
PCfun(angles = unlist(mu_diff_gl),
      col = 'darkgreen',
      shrink = 1.5,
      title = 'Green Dim - Green Bright')
# PlotCI_vM(mod_CI_tb['green_hl',],
#           col = 'darkgreen',
#           radius = 1+0.3)
PlotCI_vM(mle_CI['diff green_hl',],
          col = 'darkgreen',
          radius = 1+0.3)

plot.circular(x = circular(unlist(mu_diff_ul),
                           units = angle_unit,
                           rot = angle_rot
                           ),
              col = 'purple',
              stack = TRUE,
              sep = 0.1,
              bins = 355/5,
              units = angle_unit,
              rotation = angle_rot,
              zero = pi/2,
              shrink = 1.5)
mtext(text = 'UV Dim - Green Bright',
      side = 1,
      line = -2)
lines(x = c(0,0),
      y = c(-1,1),
      col = 'gray')
PlotMV_circMLE(mod_par = mod_diff_par['uvg_hl',],
               col1 = 'purple')
PlotCI_vM(mod_CI_tb['uvg_hl',],
          col = 'purple',
          radius = 1+0.3)

PCfun(angles = unlist(mu_diff_uh),
      col = 'gray40',
      shrink = 1.5,
      title = 'UV Bright - Green Bright')
PlotCI_vM(mod_CI_tb['uvg_h',],
          col = 'gray40',
          radius = 1+0.3)
#according to model comparison, this is slightly better
#But this is not truly bimodal, kappa2 = 0
  # plot.circular(x = circular(unlist(mu_diff_uh),
  #                            units = angle_unit,
  #                            rot = angle_rot
  #                            ),
  #               col = 'gray40',
  #               stack = TRUE,
  #               sep = 0.1,
  #               bins = 355/5,
  #               units = angle_unit,
  #               rotation = angle_rot,
  #               zero = pi/2,
  #               shrink = 1.5)
  # mtext(text = 'UV Bright - Green Bright',
  #       side = 1,
  #       line = -2)
  # lines(x = c(0,0),
  #       y = c(-1,1),
  #       col = 'gray')
  # PlotMV_circMLE(mod_par = bim_diff_par['uvg_h',],
  #                col1 = 'gray40')
  # PlotCI_vM(bim_CI_tb['uvg_h',],
  #           col = 'gray40',
  #           radius = 1+0.3)



plot.circular(x = circular(unlist(mu_diff_gul),
                           units = angle_unit,
                           rot = angle_rot
                           ),
              col = 'gray25',
              stack = TRUE,
              sep = 0.1,
              bins = 355/5,
              units = angle_unit,
              rotation = angle_rot,
              zero = pi/2,
              shrink = 1.5)
mtext(text = 'UV Dim - Green Dim',
      side = 1,
      line = -2)
lines(x = c(0,0),
      y = c(-1,1),
      col = 'gray')
PlotMV_circMLE(mod_par = mod_diff_par['uvg_l',],
               col1 = 'gray25')
PlotCI_vM(mod_CI_tb['uvg_l',],
          col = 'gray25',
          radius = 1+0.3)
```

Here we see that the confidence interval for the difference between the green conditions overlaps with 0°, and the one for the two bright conditions almost does (the one sided CI does, but looks odd due it its asymmetry).
To confirm whether these models better explain the data than the null hypothesis of the same mean for each condition (no difference), we can perform a likelihood-ratio test.

## Likelihood ratio tests

For each condition, compare the models for same mean, different mean and multiple means.
First perform all tests separately, then collect them together and finally adjust for multiple tests before reporting.

```{r Perform LR tests}
#TODO add turn angles
mod_details= lapply(X = names(pair_diffs_lst),
                     FUN = CollectDetails,
                     ll = all_res,
                     bimod = TRUE)
names(mod_details) = names(pair_diffs_lst)

#set up tests for hypotheses
lr_tests = c('pairs_diff_zero', #is the turn different from zero?
             'pairs_multi_diff', #is the turn in one or two directions?
             'pairs_multi_zero' #are the two directions different from zero?
             )#multiple tests

#Perform collect likelihood ratio tests for each hypothesis
all_lr_results_lst = lapply(X =mod_details,
                           FUN = LR_calc_lst,
                           tst = lr_tests)
all_lr_results = do.call(what = rbind,
                          args = all_lr_results_lst
                          )
all_lr_results = within(data.frame(all_lr_results),
                        {
                            comparison = c( sapply(X = names(mod_details), 
                            FUN = rep,
                            times = length(lr_tests)
                            ) )
                            test = rep(lr_tests, 
                                      times = length(names(mod_details))
                            ) 
                        }
                      )
# lapply(
#        all_lr_results_lst,
#        FUN = print,
#        digits = 3
#        )

#add labels to interpret each test according to p value and sign of comparison
all_lr_results = within(all_lr_results,
                     {
                       result = mapply(tst = lr_tests,
                                       d0 = unlist(dev0),
                                       d1 = unlist(dev1),
                                       pa = p_adjusted,
                                       FUN = H1label
                       )
                     }
)
print_lr_results = all_lr_results
print_lr_results = within(print_lr_results,
                          {
                            rm(dev0,dev1, p)
                          test = sub(sub(sub(test, 
                                     pattern = '_', 
                                     replacement = ' '),
                                     pattern = '_', 
                                     replacement = ' '),
                                     pattern = 'pairs ',
                                     replacement = '')
                          mu1 = round(all_res$mu1, 1)
                          }
                          )
print_lr_results = within(print_lr_results,
                           {mu2 = round(all_res$mu2, 1)}
                          )
rownames(print_lr_results) = with(print_lr_results, paste(comparison, test))#better to keep them for the html
#Round numeric columns
# print_lr_results = apply(print_lr_results, MARGIN = 2, FUN = unlist)
num_col = sapply(X = print_lr_results, FUN = is.numeric)
print_lr_results[,num_col] = apply(print_lr_results[,num_col], MARGIN = 2, FUN = round, digits = 3)
#Print the results
print(print_lr_results,
      digits = 3)

```

These tests confirm, that there is no significant difference between the two green conditions, a significant 13° left turn between the bright conditions, and significant bimodal differences between the two UV conditions and the two dim conditions, driven by left and right turns for the dim UV condition.
