---
title: "GUV_model_figure"
author: "James Foster"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
    fig_caption: true
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Details

**DESCRIPTION**

Load and plot circular GLMs to describe all four stimulus conditions.

**INPUTS**

`GUV_functions.R`

**OUTPUTS**

Plotted estimates.

**REFERENCES**

**Experimental approach**

Edrich W, Neumeyer C & von Helversen, O.
(1979) “Anti-sun orientation” of bees with regard to a field of ultraviolet light.
*J. Comp. Physiol.* 134, 151–157.

Rossel, S.
& Wehner, R.
(1984).
Celestial orientation in bees: the use of spectral cues.
*J. Comp. Physiol. A* 155, 605–613.

**modelling methods**

Sayin S, Couzin-Fuchs E, Petelski I, Günzel Y, Salahshour M, Lee CY, Graving JM, Li L, Deussen O, Sword GA, et al. (2025) The behavioral mechanisms governing collective motion in swarming locusts.
*Science* 387(6737):995–791

Capretto, T., Piho, C., Kumar, R., Westfall, J., Yarkoni, T., & Martin, O. (2022)
`Bambi`: A simple interface for fitting Bayesian linear models in Python.
*Journal of Statistical Software* 103(15), 1-29. doi:10.18637/jss.v103.i15

# To Do
- Load Bambi model
- Plot predictions
- Plot contrasts

# Motivation


# Set up workspace

To extract these models, we need to load the appropriate functions.
Make sure that the `circular`, `reticulate`, `posterior` and `bayesplot`,  are installed, and 
(on Windows) that you have installed the latest version of [Rtools](https://cran.r-project.org/bin/windows/Rtools/).
Here we first set up a `reticulate` Python environment with the appropriate versions of each package, then reload reticulate to ensure that those packages are available.

```{r Load functions and packages}
source('GUV_functions.R')
# require(reticulate)
p311=reticulate::install_python(version = '3.11')
reticulate::virtualenv_create(
  envname = "GUV-env",
  python = p311)
reticulate::py_install(
  envname = "GUV-env",
  packages = 
            c("numpy==1.26.4",
               "pandas==2.1.4",
               "scipy==1.11.4",
               "arviz==0.17.1",
              "h5py==3.11.0")
            )
require(reticulate)
use_virtualenv("GUV-env",
               required = TRUE)
reticulate::py_config()
require(posterior)
require(bayesplot)
options(posterior.num_args = list(sigfig = 3))#for ease of reading
```

Set up some colours for plotting.

```{r Set up plot colours}
col_kappa = '#1E78B5'#colour for kappa parameter
col_rho = '#F08024' #colour for mean vector length
col_sd = '#E74A29' #colour for SD and mean angle
col_sd2 = '#E57461' #colour for other SD heuristics
col_pdf = adjustcolor(col = '#21A885', # colour for probability density
                      alpha.f = 0.7)
col_obs = '#3E1F51' #colour for control observations
col_treat = '#006400' # colour for treatment observations
```

# Prepare data

Then load the data that will be used.
This is our dance data, reorganised by `GUV_organisedata.R` as `colour_dance_reorg.csv`.

```{r Load data}
#Dance angle data
cd = read.table(file = '1Data/colour_dance_reorg.csv', 
                header = T, 
                sep  = ',')
head(cd)
```

## Format data

Ensure that each variable has the correct data format for subsequent analysis.

```{r Data formatting}
cd = within(cd,
            {
              ID = as.factor(ID) # beedance identifier as a factor
              date = as.factor(date) # date as a factor
              signed_angle = Mod360.180(bearing)  # bearing between -180 and 180
              angle = circular(rad(signed_angle),# bearing between -pi and pi
                               rotation = 'clock') # circular format suppresses later warnings
            }
)
u_id = with(cd, unique(ID)) # unique beedances
length(u_id)#169 beedances
```
```{r Shorten variable names}
cd = within(cd,
            {
              BR = brightn
              CL = colour
              DT = date
              RN = run
            }
)
```

## Identify dances for both conditions
Identify dances in which all four stimulus conditions were used.

```{r both conditions}
uvc = sapply(u_id,
              FUN = function(id, dt)
                {
                with(subset(dt,
                            ID %in% id & 
                              colour %in% 'u'),
                     {
                       length(unique(brightn))
                       }
                     )
              },
              dt = cd)

#There should be two brightness levels for UV
uv_ids = u_id[uvc == 2]
length(uv_ids)#most individuals
#extract just those individuals

#find data for full condition individuals
cd_uv = subset(cd, ID %in% uv_ids)
```

# Load Bambi model and extract draws for each parameter

```{r Load the model}
#Import arviz for model handling
az = reticulate::import("arviz")
# Load the NetCDF file
idata = az$from_netcdf("2Results/GUV_unwrap_model.nc")
# Convert to data frame
posterior_df = reticulate::py_to_r(x =                                         
                                     idata$posterior$to_dataframe()
                                   )
```

Set up metadata about draws for sorting.

```{r Identify individuals and chains}
# Initial Metadata -----------------------------------------------------

n_ind = 169
n_chains = 4
obs_per_chain = nrow(posterior_df) /
                    (n_ind * n_chains) # Calculated automatically

# Individual changes every row (1,2...169, 1,2...169)
posterior_df = within(posterior_df,
                      {
                Individual_ID = rep(x = 1:n_ind, 
                         times = n_chains * obs_per_chain)
    
    # Iteration changes every 169 rows (1,1... (169 times), 2,2... (169 times))
                .iteration    = rep(x = rep(x = 1:obs_per_chain, 
                                 each = n_ind), 
                         times = n_chains)

# Chain changes every (169 * obs_per_chain) rows
                .chain        = rep(x = 1:n_chains, 
                                    each = obs_per_chain * n_ind)
                      }
)
# Define Parameter Groupings
global_params = c(
  "Intercept", "B", "C", "B:C", 
  "kappa_Intercept", "kappa_B", "kappa_C", "kappa_B:C",
  "1|Individual_kappa", "B|Individual_kappa", "C|Individual_kappa", "B:C|Individual_kappa",
  "kappa_1|Individual_sigma", "kappa_B|Individual_sigma", 
  "kappa_C|Individual_sigma", "kappa_B:C|Individual_sigma",
  "kappa_1|Individual_offset", "kappa_B|Individual_offset", 
  "kappa_C|Individual_offset", "kappa_B:C|Individual_offset"
)

ind_params = c(
  "1|Individual", "B|Individual", "C|Individual", "B:C|Individual",
  "kappa_1|Individual", "kappa_B|Individual", "kappa_C|Individual", "kappa_B:C|Individual"
)

```

```{r Reorganise draws by individual}

# Reshape posterior draws -------------------------------------------------

# Reshape to Wide Format
list_of_inds = split(x = posterior_df, 
                     f = posterior_df$Individual_ID)

# Start wide df with metadata and population-level parameters
existing_globals = intersect(x = global_params,
                             y = colnames(posterior_df))
# take the first individual's data as a template for the wide data frame
draws_wide  =  list_of_inds[[1]][, c(".chain", ".iteration", existing_globals)]

# Pivot individual-level parameters into columns
for(param in ind_params)
  {
  if(param %in% colnames(posterior_df))
    {
    clean_base = gsub("\\|", "_", param)
    for(i in 1:n_ind) {
      draws_wide[[paste0(clean_base, "_Ind_", i)]] = list_of_inds[[i]][[param]]
    }
  }
}

# Export to draws dataframe
draws_final = posterior::as_draws_df(draws_wide)
```


# Inspect model estimates

This model identifies both a population level bias (near 0°) and individual biases
 with stable parameter estimates on the 'unwrapped' (circular modulo) scale.
 
```{r Inspect unimodal change model}

mcmc_trace(draws_final,
           pars = c('Intercept', 'B', 'C', 'B:C'),
           transformations = unwrap_circular_deg)
mcmc_trace(draws_final,
           pars = paste0("kappa_",
                         c('Intercept', 'B', 'C', 'B:C')) )
mcmc_trace(draws_final,
           pars = paste0( c('1', 'B', 'C', 'B:C'),
                          '|Individual_kappa') )
mcmc_trace(draws_final,
           pars = paste0('kappa_',
                         c('1', 'B', 'C', 'B:C'),
                         '|Individual_sigma') )
```



```{r Inspect individual means unimodal change model}
plt_panels = 9
#individual but they overlap when unwrapped
for(id in 0:(ceiling(length(u_id)/plt_panels)-1))
{
  # Calculate the specific range of indices for this page
  start_idx = (id * plt_panels) + 1
  end_idx   = min((id + 1) * plt_panels, length(u_id))
plt = mcmc_trace(draws_final,
           pars = paste0("1_Individual_Ind_",
                         start_idx:end_idx),
           facet_args = list(ncol = 3,
                             nrow = 3),
           transformations = unwrap_circular_deg
           )
print(plt)
}
```

Calculating the $\hat{R}$ values for unwrapped estimates for $\mu_i$ reveals that 
all estimates converged well ($\hat{R}<1.01$).

```{r summarise fixef with arviz}
azsum_fixed = az$summary(idata,
           var_names=c("Intercept", #slightly right shifted, should converge to around 0.26
               "B", 
                "C",
                "B:C",
                "kappa_Intercept",
               "kappa_B",
                "kappa_C",
                "kappa_B:C"
                ),
    circ_var_names=c("Intercept",
                    "B",
                    "C",
                    "B:C"
                    )
)
print(azsum_fixed[, c(1, 3, 4, 7, 9)], digits = 3)
```

```{r summarise ranef with arviz}
azsum_ranef = az$summary(idata,
           var_names=c( "1|Individual_kappa",
        "B|Individual_kappa",
        "C|Individual_kappa",
        "B:C|Individual_kappa",
        "kappa_1|Individual_sigma",
        "kappa_B|Individual_sigma",
        "kappa_C|Individual_sigma",
        "kappa_B:C|Individual_sigma"
                )
)
print(azsum_ranef[, c(1, 3, 4, 7, 9)], digits = 3)
```

## Extract predictions

```{r set up functions to summarise draws}

#A function to find the circular median in (-pi, pi)
MedCDraws = function(x,
                     templ = 'none', #assumes radians
                     medn = 1) #take the 1st median
{
  median.circular( circular(x = mod_circular(x),
                            template = templ)
                                           )[medn]
}
QCDraws = function(x,
                   probs = c(0.025, 0.5, 0.975),
                     templ = 'none',
                     medn = 1)
{
  quantile.circular( circular(x = mod_circular(x),
                            template = templ),
                     probs = probs
                                           )
}
```

```{r Extract individual predictions}
# Setup containers for summary statistics
u_id_num <- 1:n_ind
mu_id_gh <- kappa_id_gh <- numeric(length(u_id_num))
mu_id_uh <- kappa_id_uh <- numeric(length(u_id_num))
mu_id_gl <- kappa_id_gl <- numeric(length(u_id_num))
mu_id_ul <- kappa_id_ul <- numeric(length(u_id_num))

# Create one Master List for all raw draws
# Structure: individual_draws[[i]]$mu$gh
individual_draws <- vector("list", length(u_id_num))

for(i in seq_along(u_id_num)) {
  
  calc <- within(draws_final, {
    # Dynamic fetch
    m_i   <- get(paste0("1_Individual_Ind_", i))
    m_bi  <- get(paste0("B_Individual_Ind_", i))
    m_ci  <- get(paste0("C_Individual_Ind_", i))
    m_bci <- get(paste0("B:C_Individual_Ind_", i))
    
    k_i   <- get(paste0("kappa_1_Individual_Ind_", i))
    k_bi  <- get(paste0("kappa_B_Individual_Ind_", i))
    k_ci  <- get(paste0("kappa_C_Individual_Ind_", i))
    k_bci <- get(paste0("kappa_B:C_Individual_Ind_", i))
    
    # Calculate Raw Vectors
    gh_mu_vec <- Intercept + m_i
    gh_ka_vec <- kappa_Intercept + k_i
    
    uh_mu_vec <- Intercept + m_i + C + m_ci
    uh_ka_vec <- kappa_Intercept + k_i + kappa_C + k_ci
    
    gl_mu_vec <- Intercept + m_i + B + m_bi
    gl_ka_vec <- kappa_Intercept + k_i + kappa_B + k_bi
    
    ul_mu_vec <- Intercept + m_i + B + C + `B:C` + m_bi + m_ci + m_bci
    ul_ka_vec <- kappa_Intercept + k_i + kappa_B + kappa_C + `kappa_B:C` + k_bi + k_ci + k_bci
  })
  
  # Save to the Nested List
  individual_draws[[i]] <- list(
    mu    = list(gh = calc$gh_mu_vec, uh = calc$uh_mu_vec, gl = calc$gl_mu_vec, ul = calc$ul_mu_vec),
    kappa = list(gh = calc$gh_ka_vec, uh = calc$uh_ka_vec, gl = calc$gl_ka_vec, ul = calc$ul_ka_vec)
  )
  
  # Store summary stats for quick table access
  mu_id_gh[i] <- MedCDraws(calc$gh_mu_vec)
  kappa_id_gh[i] <- median(calc$gh_ka_vec)
  
  mu_id_uh[i] <- MedCDraws(calc$uh_mu_vec)
  kappa_id_uh[i] <- median(calc$uh_ka_vec)
  
  mu_id_gl[i] <- MedCDraws(calc$gl_mu_vec)
  kappa_id_gl[i] <- median(calc$gl_ka_vec)
  
  mu_id_ul[i] <- MedCDraws(calc$ul_mu_vec)
  kappa_id_ul[i] <- median(calc$ul_ka_vec)
}
```

```{r Plot the unimodal model}



par(pty = 's')
par(mar = c(0,0,0,0),
    mfrow = c(4,4) )
for(i in 1:length(u_id) )
# for(i in 20+1:50 )
{
  
  # Construct the exact column names
  mu_i   = paste0("1_Individual_Ind_", i)
  mu_bi  = paste0("B_Individual_Ind_", i)
  mu_ci  = paste0("C_Individual_Ind_", i)
  mu_bci = paste0("B:C_Individual_Ind_", i)
  
  ka_i   = paste0("kappa_1_Individual_Ind_", i)
  ka_bi  = paste0("kappa_B_Individual_Ind_", i)
  ka_ci  = paste0("kappa_C_Individual_Ind_", i)
  ka_bci = paste0("kappa_B:C_Individual_Ind_", i)
  
  #green high
  with(subset(cd, 
              ID %in% u_id[i] &
                CL %in% 'g' & BR %in% 'h'),
  {
    PCfun(angles = bearing,
         col = 'green3',
         plot_rho = FALSE,
         title = u_id[i])
    }
  )
  Draws2Cont(individual_draws[[i]],
           x_string = paste0('sin(mu$gh) * A1(softplus(kappa$gh))'),
           y_string = paste0('cos(mu$gh) * A1(softplus(kappa$gh))')
  )
  
  arrows.circular(
    x = circular(mu_id_gh[i],
                 units = 'radians',
                 rotation = 'clock',
                 zero = pi/2),
    y = A1(softplus(kappa_id_gh[i])), # Ensure softplus is applied here
    lwd = 5,
    length = 0.1/1.25,
    col = adjustcolor('green3', alpha.f = 0.8)
  )

  #UV high
  with(subset(cd,
              ID %in% u_id[i] &
                CL %in% 'u' & BR %in% 'h'),
  {
    PCfun(angles = bearing,
         col = 'magenta3',
         plot_rho = FALSE,
         title = u_id[i])
    }
  )
   Draws2Cont(individual_draws[[i]],
           x_string = paste0('sin(mu$uh) * A1(softplus(kappa$uh))'),
           y_string = paste0('cos(mu$uh) * A1(softplus(kappa$uh))')
  )
  arrows.circular(x = circular(x = mu_id_uh[i],
                               units = 'radians',
                              rotation = 'clock',
                              zero = pi/2),
                 y = A1(kappa_id_uh[i]),
                 lwd = 5,
                 length = 0.1/1.25,
                 col = adjustcolor('magenta3', alpha.f = 200/255))

  #green low
  with(subset(cd,
              ID %in% u_id[i] &
                CL %in% 'g' & BR %in% 'l'),
  {
    PCfun(angles = bearing,
         col = 'green4',
         plot_rho = FALSE,
         title = u_id[i])
    }
  )
   Draws2Cont(individual_draws[[i]],
           x_string = paste0('sin(mu$gl) * A1(softplus(kappa$gl))'),
           y_string = paste0('cos(mu$gl) * A1(softplus(kappa$gl))')
  )
  arrows.circular(x = circular(x = mu_id_gl[i],
                               units = 'radians',
                              rotation = 'clock',
                              zero = pi/2),
                 y = A1(kappa_id_gl[i]),
                 lwd = 5,
                 length = 0.1/1.25,
                 col = adjustcolor('green4', alpha.f = 200/255))
  # 
  #UV low
  with(subset(cd,
              ID %in% u_id[i] &
                CL %in% 'u' & BR %in% 'l'),
  {
    PCfun(angles = bearing,
         col = 'purple',
         plot_rho = FALSE,
         title = u_id[i])
    }
  )
  Draws2Cont(individual_draws[[i]],
           x_string = paste0('sin(mu$ul) * A1(softplus(kappa$ul))'),
           y_string = paste0('cos(mu$ul) * A1(softplus(kappa$ul))')
  )
  arrows.circular(x = circular(x = mu_id_ul[i],
                               units = 'radians',
                              rotation = 'clock',
                              zero = pi/2),
                 y = A1(kappa_id_ul[i]),
                 lwd = 5,
                 length = 0.1/1.25,
                 col = adjustcolor('purple', alpha.f = 200/255))
}

```


## Plot population level predictions


```{r Calculate estimates of means and changes}

#calculate mean vectors
mean_vectors = aggregate(angle~ID*brightn*colour,#+sun_az+sun_el_rad, # N.B. Including sun azimuth drops some cases without a time stamp
                         data = cd,
                         FUN = rho.circular
)
#correct names
mean_vectors = within(mean_vectors,
                      {mean_vector = angle; rm(angle)} # angle now indicates a mean vector, not an angle
)
#add kappa
mle_estimates = aggregate(angle~ID*brightn*colour,#+sun_az+sun_el,
                          data = cd,
                          FUN = MLE_est
)
#Rayleigh tests
rayleigh_tests = aggregate(angle~ID*brightn*colour,#+sun_az+sun_el,
                          data = cd,
                          FUN = Rayleigh_p
                          )
#add to the summary table and 
#calculate inverse softplus kappa
mean_vectors = within(mean_vectors,
                      {
                        mu = deg(mle_estimates$angle[,'mu'])
                        kappa = mle_estimates$angle[,'kappa']
                        iskappa  = inv_softplus(kappa)
                      }
)

luc = sapply(u_id,
              FUN = IndCond,
              dt = cd)

#Most individuals that made it to bright green 
full_ids = u_id[luc == 4]

#find pair conditions
#green
grc = sapply(u_id,
              FUN = function(id, dt)
                {
                with(subset(dt,
                            ID %in% id & 
                              colour %in% 'g'),
                     {
                       length(unique(brightn))
                       }
                     )
              },
              dt = cd)
#bright
brc = sapply(u_id,
              FUN = function(id, dt)
                {
                with(subset(dt,
                            ID %in% id & 
                              brightn %in% 'h'),
                     {
                       length(unique(colour))
                       }
                     )
              },
              dt = cd)
#UV
uvc = sapply(u_id,
              FUN = function(id, dt)
                {
                with(subset(dt,
                            ID %in% id & 
                              colour %in% 'u'),
                     {
                       length(unique(brightn))
                       }
                     )
              },
              dt = cd)
#dim
dic = sapply(u_id,
              FUN = function(id, dt)
                {
                with(subset(dt,
                            ID %in% id & 
                              brightn %in% 'l'),
                     {
                       length(unique(colour))
                       }
                     )
              },
              dt = cd)

#There should be two brightness levels for green
gr_ids = u_id[grc == 2]
#There should be two colour levels for bright
br_ids = u_id[brc == 2]
#There should be two brightness levels for UV
uv_ids = u_id[uvc == 2]
#There should be two colour levels for dim
di_ids = u_id[dic == 2]

#condition specific
gh_ids = with(subset(cd,
                     subset = CL %in% 'g' & 
                       BR %in% 'h'),
              unique(ID)
              )
uh_ids = with(subset(cd,
                     subset = CL %in% 'u' & 
                       BR %in% 'h'),
              unique(ID)
              )
gl_ids = with(subset(cd,
                     subset = CL %in% 'g' & 
                       BR %in% 'l'),
              unique(ID)
              )
ul_ids = with(subset(cd,
                     subset = CL %in% 'u' & 
                       BR %in% 'l'),
              unique(ID)
              )


mu_diff_gl = sapply(X = gr_ids,
                    FUN = MuDiff,
                    dt = mean_vectors,
                    cl = 'g',
                    br = 'l')
mu_diff_uh = sapply(X = br_ids,
                    FUN = MuDiff,
                    dt = mean_vectors,
                    cl = 'u',
                    br = 'h')
mu_diff_ul = sapply(X = uv_ids,
                    FUN = MuDiff,
                    dt = mean_vectors,
                    cl = 'u',
                    br = 'l')
mu_diff_ugl = sapply(X = di_ids,
                    FUN = MuDiff,
                    dt = mean_vectors,
                     cl = 'u',
                    br = 'l',
                    ref_cl = 'g',
                    ref_br = 'h')

```

```{r Extract population level predictions}
par(pty = 's')
# Population mean for green bright (Baseline: Intercept)
par(mfrow = c(1,4), mar = c(0,0,0,0))

with(subset(mean_vectors, colour %in% 'g' & brightn %in% 'h'), {
  PCfun(angles = mu, col = 'green3', shrink = 2, plot_rho = FALSE)
})

for(i in which(u_id %in% gh_ids)) {
  arrows.circular(x = circular(x = mu_id_gh[i], units = 'radians', rotation = 'clock', zero = pi/2),
                  y = A1(kappa_id_gh[i]), lwd = 0.25, length = 0,
                  col = adjustcolor('green3', alpha.f = 200/255))
}

Draws2Cont(draws = draws_final,
           x_string = 'sin(Intercept) * A1(softplus(kappa_Intercept))',
           y_string = 'cos(Intercept) * A1(softplus(kappa_Intercept))')

with(draws_final,
     arrows.circular(x = circular(x = MedCDraws(Intercept), 
                                  units = 'radians', 
                                  rotation = 'clock', zero = pi/2),
                     y = A1(median(softplus(kappa_Intercept))),
                     lwd = 5, length = 0.1/1.25,
                     col = adjustcolor('green3', alpha.f = 200/255, offset = c(rep(-0.2,3),0) )
     )
)


# pop mean for UV bright (Intercept + C)
with(subset(mean_vectors, colour %in% 'u' & brightn %in% 'h'), {
  PCfun(angles = mu, col = 'magenta3', shrink = 2, plot_rho = FALSE)
})

for(i in which(u_id %in% uh_ids) )
{

  arrows.circular(x = circular(x = mu_id_uh[i],
                               units = 'radians',
                              rotation = 'clock',
                              zero = pi/2),
                 y = A1(kappa_id_uh[i]),
                 lwd = 0.25,
                 length = 0,
                 col = adjustcolor('magenta3', alpha.f = 200/255))
}

Draws2Cont(draws = draws_final,
           x_string = 'sin(Intercept + C) * A1(softplus(kappa_Intercept + kappa_C))',
           y_string = 'cos(Intercept + C) * A1(softplus(kappa_Intercept + kappa_C))')

with(draws_final,
     arrows.circular(x = circular(x = MedCDraws(Intercept + C), units = 'radians', rotation = 'clock', zero = pi/2),
                     y = A1(median(softplus(kappa_Intercept + kappa_C))),
                     lwd = 5, length = 0.1/1.25, col = adjustcolor('magenta3', alpha.f = 200/255, offset = c(rep(-0.2,3),0)))
)

# pop mean for green dim (Intercept + B)
with(subset(mean_vectors, colour %in% 'g' & brightn %in% 'l'), {
  PCfun(angles = mu, col = 'darkgreen', shrink = 2, plot_rho = FALSE)
})


for(i in which(u_id %in% gl_ids) )
{

  arrows.circular(x = circular(x = mu_id_gl[i],
                               units = 'radians',
                              rotation = 'clock',
                              zero = pi/2),
                 y = A1(kappa_id_gl[i]),
                 lwd = 0.25,
                 length = 0,
                 col = adjustcolor('darkgreen', alpha.f = 200/255 ))
}

Draws2Cont(draws = draws_final,
           x_string = 'sin(Intercept + B) * A1(softplus(kappa_Intercept + kappa_B))',
           y_string = 'cos(Intercept + B) * A1(softplus(kappa_Intercept + kappa_B))')

with(draws_final,
     arrows.circular(x = circular(x = MedCDraws(Intercept + B), units = 'radians', rotation = 'clock', zero = pi/2),
                     y = A1(median(softplus(kappa_Intercept + kappa_B))),
                     lwd = 5, length = 0.1/1.25, col = adjustcolor('darkgreen', alpha.f = 200/255, offset = c(rep(-0.2,3),0)))
)

# pop mean for UV dim (Intercept + B + C + BC)
with(subset(mean_vectors, colour %in% 'u' & brightn %in% 'l'), {
  PCfun(angles = mu, col = 'purple', shrink = 2, plot_rho = FALSE)
})


for(i in which(u_id %in% ul_ids) )
{

  arrows.circular(x = circular(x = mu_id_ul[i],
                               units = 'radians',
                              rotation = 'clock',
                              zero = pi/2),
                 y = A1(kappa_id_ul[i]),
                 lwd = 0.25,
                 length = 0,
                 col = adjustcolor('purple', alpha.f = 200/255)
  )
}

Draws2Cont(draws = draws_final,
           x_string = 'sin(Intercept + B + C + `B:C`) * A1(softplus(kappa_Intercept + kappa_B + kappa_C + `kappa_B:C`))',
           y_string = 'cos(Intercept + B + C + `B:C`) * A1(softplus(kappa_Intercept + kappa_B + kappa_C + `kappa_B:C`))')

with(draws_final,
     arrows.circular(x = circular(x = MedCDraws(Intercept + B + C + `B:C`), units = 'radians', rotation = 'clock', zero = pi/2),
                     y = A1(median(softplus(kappa_Intercept + kappa_B + kappa_C + `kappa_B:C`))),
                     lwd = 5, length = 0.1/1.25, col = adjustcolor('purple', alpha.f = 200/255, offset = c(rep(-0.2,3),0)))
)
```

```{r relative pop level effects}
par(pty = 's')
# Population mean for green bright (Baseline: Intercept)
par(mfrow = c(1,4), mar = c(0,0,0,0))

# --- Effect of UV ---
PCfun(angles = unlist(mu_diff_uh),
      col = 'gray40',
      shrink = 2,
      plot_rho = FALSE)

for(i in which(u_id %in% uv_ids)) {
  # UV high
  arrows.circular(x = circular(x = mu_id_uh[i] - mu_id_gh[i],
                               units = 'radians',
                               rotation = 'clock',
                               zero = pi/2),
                  y = A1(kappa_id_uh[i]),
                  lwd = 0.5,
                  length = 0,
                  col = adjustcolor('magenta3', alpha.f = 200/255))
}

Draws2Cont(draws = draws_final,
           x_string = 'sin(C) * A1(softplus(kappa_C))',
           y_string = 'cos(C) * A1(softplus(kappa_C))')

with(draws_final,
     arrows.circular(x = circular(x = MedCDraws(C),
                                  units = 'radians',
                                  rotation = 'clock',
                                  zero = pi/2),
     y = A1(median(softplus(kappa_C))),
     lwd = 5,
     length = 0.1/1.25,
     col = adjustcolor('gray40', alpha.f = 200/255))
)

with(draws_final,
     {
       round(deg(
         quantile(mod_circular(C), probs = c(0.5, 0.025, 0.975))
       ))
     }
)

# --- Effect of dim green ---
PCfun(angles = unlist(mu_diff_gl),
      col = 'darkgreen',
      shrink = 2,
      plot_rho = FALSE)

# green low
for(i in which(u_id %in% gr_ids)) {
  arrows.circular(x = circular(x = mu_id_gl[i] - mu_id_gh[i],
                               units = 'radians',
                               rotation = 'clock',
                               zero = pi/2),
                  y = A1(kappa_id_gl[i]),
                  lwd = 0.25,
                  length = 0,
                  col = adjustcolor('green4', alpha.f = 200/255))
}

Draws2Cont(draws = draws_final,
           x_string = 'sin(B) * A1(softplus(kappa_B))',
           y_string = 'cos(B) * A1(softplus(kappa_B))')

with(draws_final,
     {
       round(deg(
         quantile(mod_circular(B), probs = c(0.5, 0.025, 0.975))
       ))
     }
)

with(draws_final,
     arrows.circular(x = circular(x = MedCDraws(B),
                                  units = 'radians',
                                  rotation = 'clock',
                                  zero = pi/2),
     y = A1(median(softplus(kappa_B))),
     lwd = 5,
     length = 0.1/1.25,
     col = adjustcolor('darkgreen', alpha.f = 200/255))
)

# --- Effect of dim UV ---
PCfun(angles = unlist(mu_diff_ul),
      col = 'purple',
      shrink = 2,
      plot_rho = FALSE)

# UV low
for(i in which(u_id %in% uv_ids)) {
  arrows.circular(x = circular(x = mu_id_ul[i] - mu_id_uh[i],
                               units = 'radians',
                               rotation = 'clock',
                               zero = pi/2),
                  y = A1(kappa_id_ul[i]),
                  lwd = 0.25,
                  length = 0,
                  col = adjustcolor('purple', alpha.f = 200/255))
}

# The difference between UV bright and UV dim is brightness and the interaction
Draws2Cont(draws = draws_final,
           x_string = 'sin(B + `B:C`) * A1(softplus(kappa_B + `kappa_B:C`))',
           y_string = 'cos(B + `B:C`) * A1(softplus(kappa_B + `kappa_B:C`))')

with(draws_final,
     {
       round(deg(
         quantile(mod_circular(B + `B:C`), probs = c(0.5, 0.025, 0.975))
       ))
     }
)

with(draws_final,
     arrows.circular(x = circular(x = MedCDraws(B + `B:C`),
                                  units = 'radians',
                                  rotation = 'clock',
                                  zero = pi/2),
     y = A1(median(softplus(kappa_B + `kappa_B:C`))),
     lwd = 5,
     length = 0.1/1.25,
     col = adjustcolor('purple', alpha.f = 200/255))
)

# --- Effect of dim UV relative to dim green ---
PCfun(angles = unlist(mu_diff_ugl),
      col = 'gray25',
      shrink = 2,
      plot_rho = FALSE)

# UV low
for(i in which(u_id %in% di_ids)) {
  arrows.circular(x = circular(x = mu_id_ul[i] - mu_id_gl[i],
                               units = 'radians',
                               rotation = 'clock',
                               zero = pi/2),
                  y = A1(kappa_id_ul[i]),
                  lwd = 0.25,
                  length = 0,
                  col = adjustcolor('gray25', alpha.f = 200/255))
}

# The difference between green dim and UV dim is colour and the interaction
Draws2Cont(draws = draws_final,
           x_string = 'sin(C + `B:C`) * A1(softplus(kappa_C + `kappa_B:C`))',
           y_string = 'cos(C + `B:C`) * A1(softplus(kappa_C + `kappa_B:C`))')

with(draws_final,
     {
       round(deg(
         quantile(mod_circular(C + `B:C`), probs = c(0.5, 0.025, 0.975))
       ))
     }
)

with(draws_final,
     arrows.circular(x = circular(x = MedCDraws(C + `B:C`),
                                  units = 'radians',
                                  rotation = 'clock',
                                  zero = pi/2),
     y = A1(median(softplus(kappa_C + `kappa_B:C`))),
     lwd = 5,
     length = 0.1/1.25,
     col = adjustcolor('gray25', alpha.f = 200/255))
)
```

```{r Kappa predictions}
#plot kappa
par(mfrow = c(1,4))

with(draws_final,
     VertHist(data = softplus(kappa_Intercept), 
              main = '\ngreen bright',
              ylab = 'kappa',
              ylim = c(0, 15),
              col = adjustcolor('green3', alpha.f = 100/255),
              axes = FALSE,
              cex.axis = 0.7))
axis(1)
axis(2, at = 0:5*5)

with(draws_final,
     VertHist(data = softplus(kappa_Intercept + kappa_C), 
              main = '\nUV bright',
              ylab = 'kappa',
              ylim = c(0, 15),
              col = adjustcolor('magenta3', alpha.f = 100/255),
              axes = FALSE,
              cex.axis = 0.7))
axis(1)
axis(2, at = 0:5*5)

with(draws_final,
     {
       round(
     quantile(softplus(kappa_Intercept + kappa_C) - 
                softplus(kappa_Intercept),
              probs = c(0.5, 0.025, 0.975)),
     digits = 2
       )
     }
)

with(draws_final,
     VertHist(data = softplus(kappa_Intercept + kappa_B), 
              main = '\ngreen dim',
              ylab = 'kappa',
              ylim = c(0, 15),
              col = adjustcolor('green4', alpha.f = 100/255),
              axes = FALSE,
              cex.axis = 0.7))
axis(1)
axis(2, at = 0:5*5)

with(draws_final,
     VertHist(data = softplus(kappa_Intercept + kappa_B  + kappa_C +  + `kappa_B:C`), 
              main = '\nUV dim',
              ylab = 'kappa',
              ylim = c(0, 15),
              col = adjustcolor('purple', alpha.f = 100/255),
              axes = FALSE,
              cex.axis = 0.7))
axis(1)
axis(2, at = 0:5*5)


with(draws_final,
     {
       round(
     quantile(softplus(kappa_Intercept + kappa_B) - 
                softplus(kappa_Intercept),
              probs = c(0.5, 0.025, 0.975)),
     digits = 2
       )
     }
)

```
``` {r Kappa effects}
# Plot difference in kappa
par(mfrow = c(1,4))

# 1. Effect of UV Brightness (UV Bright - Green Bright)
with(draws_final,
     VertHist(data = softplus(kappa_Intercept + kappa_C) - softplus(kappa_Intercept), 
              main = '\nDiff: UV bright',
              ylab = 'diff kappa',
              ylim = c(-10, 10),
              col = adjustcolor('magenta3', alpha.f = 100/255),
              axes = FALSE,
              cex.axis = 0.7))
axis(1)
axis(2, at = -2:2*5)
abline(h = 0, lty = 2)

# 2. Effect of Dimming Green (Green Dim - Green Bright)
with(draws_final,
     VertHist(data = softplus(kappa_Intercept + kappa_B) - softplus(kappa_Intercept), 
              main = '\nDiff: green dim',
              ylab = 'diff kappa',
              ylim = c(-10, 10),
              col = adjustcolor('green4', alpha.f = 100/255),
              axes = FALSE,
              cex.axis = 0.7))
axis(1)
axis(2, at = -2:2*5)
abline(h = 0, lty = 2)

# 3. Effect of Dimming UV (UV Dim - UV Bright)
with(draws_final,
     VertHist(data = softplus(kappa_Intercept + kappa_B + kappa_C + `kappa_B:C`) - 
                     softplus(kappa_Intercept + kappa_C), 
              main = '\nDiff: dim UV',
              ylab = 'diff kappa',
              ylim = c(-10, 10),
              col = adjustcolor('purple', alpha.f = 100/255),
              axes = FALSE,
              cex.axis = 0.7))
axis(1)
axis(2, at = -2:2*5)
abline(h = 0, lty = 2)

# 4. Interaction Effect (The "Difference in Differences")
with(draws_final,
     VertHist(data = softplus(kappa_B + `kappa_B:C`) - softplus(kappa_B), 
              main = '\nDim UV vs dim green',
              ylab = 'diff kappa',
              ylim = c(-10, 10),
              col = adjustcolor('gray40', alpha.f = 100/255),
              axes = FALSE,
              cex.axis = 0.7))
axis(1)
axis(2, at = -2:2*5)
abline(h = 0, lty = 2)

# Summary Stats for the UV Dimming Effect
with(draws_final,
     {
       round(
         quantile(softplus(kappa_Intercept + kappa_B + kappa_C + `kappa_B:C`) - 
                  softplus(kappa_Intercept + kappa_C), 
                  probs = c(0.5, 0.025, 0.975)),
         digits = 2
       )
     }
)
```

```{r Kappamu effects}
#plot kappa
par(mfrow = c(1,4))

with(draws_final,
     VertHist(data = softplus(`1|Individual_kappa`), 
              main = '\ngreen bright',
              ylab = 'kappamu',
              ylim = c(0, 15),
              col = adjustcolor('green3', alpha.f = 100/255),
              axes = FALSE,
              cex.axis = 0.7))
axis(1)
axis(2, at = 0:5*5)

with(draws_final,
     VertHist(data = softplus(`C|Individual_kappa`), 
              main = '\nUV effect',
              ylab = 'kappamu',
              ylim = c(0, 50),
              col = adjustcolor('gray50', alpha.f = 100/255),
              axes = FALSE,
              cex.axis = 0.7))
axis(1)
axis(2, at = 0:5*10)

with(draws_final,
     VertHist(data = softplus(`B|Individual_kappa`), 
              main = '\ndim effect',
              ylab = 'kappamu',
              ylim = c(0, 50),
              col = adjustcolor('darkgreen', alpha.f = 100/255),
              axes = FALSE,
              cex.axis = 0.7))
axis(1)
axis(2, at = 0:5*10)

with(draws_final,
     VertHist(data = softplus(`B:C|Individual_kappa`), 
              main = '\ndim uv effect',
              ylab = 'kappamu',
              ylim = c(0, 50),
              col = adjustcolor('purple', alpha.f = 100/255),
              axes = FALSE,
              cex.axis = 0.7))
axis(1)
axis(2, at = 0:5*10)
```

```{r Kappamu conditions}

# 1. Reconstruct matrices: Rows = Draws, Cols = Individuals
n_draws <- length(individual_draws[[1]]$mu$gh)
n_ind   <- length(individual_draws)

# Helper to extract a condition into a Draw-by-Individual matrix
get_mu_matrix <- function(cond) {
  sapply(1:n_ind, function(i) individual_draws[[i]]$mu[[cond]])
}

mu_mat_gh <- get_mu_matrix("gh")
mu_mat_uh <- get_mu_matrix("uh")
mu_mat_gl <- get_mu_matrix("gl")
mu_mat_ul <- get_mu_matrix("ul")

# 2. Function to calculate MLE kappa for a single row of mu draws
calc_kappa_mle <- function(mu_row) {
  # mle.vonmises returns an object; we just want the kappa estimate
  as.numeric(mle.vonmises(circular(mu_row, units='radians'))$kappa)
}

# 3. Apply across all draws (rows) to get the posterior of kappamu
post_kappamu_gh <- apply(mu_mat_gh, 1, calc_kappa_mle)
post_kappamu_uh <- apply(mu_mat_uh, 1, calc_kappa_mle)
post_kappamu_gl <- apply(mu_mat_gl, 1, calc_kappa_mle)
post_kappamu_ul <- apply(mu_mat_ul, 1, calc_kappa_mle)

# 4. Plotting
par(mfrow = c(1,4))

# Green Bright
VertHist(data = post_kappamu_gh, 
         main = '\ngreen bright', ylab = 'kappamu (MLE)',
         ylim = c(0, 5), col = adjustcolor('green3', alpha.f = 100/255),
         axes = FALSE, cex.axis = 0.7)
axis(1); axis(2, at = 0:5)

# UV Bright
VertHist(data = post_kappamu_uh, 
         main = '\nUV bright', ylab = 'kappamu (MLE)',
         ylim = c(0, 5), col = adjustcolor('magenta3', alpha.f = 100/255),
         axes = FALSE, cex.axis = 0.7)
axis(1); axis(2, at = 0:5)

# Green Dim
VertHist(data = post_kappamu_gl, 
         main = '\ngreen dim', ylab = 'kappamu (MLE)',
         ylim = c(0, 5), col = adjustcolor('green4', alpha.f = 100/255),
         axes = FALSE, cex.axis = 0.7)
axis(1); axis(2, at = 0:5)

# UV Dim
VertHist(data = post_kappamu_ul, 
         main = '\nUV dim', ylab = 'kappamu (MLE)',
         ylim = c(0, 5), col = adjustcolor('purple', alpha.f = 100/255),
         axes = FALSE, cex.axis = 0.7)
axis(1); axis(2, at = 0:5)
```
```{r kappamu condition differences}
# 1. Calculate Differences
diff_uv_bright <- post_kappamu_uh - post_kappamu_gh
diff_green_dim <- post_kappamu_gl - post_kappamu_gh
diff_uv_dim    <- post_kappamu_ul - post_kappamu_uh
diff_interaction <- diff_uv_dim - diff_green_dim

# 2. Plotting
par(mfrow = c(1,4))

# Plot 1: Effect of UV Brightness
VertHist(data = diff_uv_bright, 
         main = '\nDiff: UV bright', ylab = 'diff kappamu',
         ylim = c(-2, 2), col = adjustcolor('magenta3', alpha.f = 100/255),
         axes = FALSE, cex.axis = 0.7)
axis(1); axis(2, at = -2:2); abline(h = 0, lty = 2)

# Plot 2: Effect of Dimming Green
VertHist(data = diff_green_dim, 
         main = '\nDiff: green dim', ylab = 'diff kappamu',
         ylim = c(-2, 2), col = adjustcolor('green4', alpha.f = 100/255),
         axes = FALSE, cex.axis = 0.7)
axis(1); axis(2, at = -2:2); abline(h = 0, lty = 2)

# Plot 3: Effect of Dimming UV
VertHist(data = diff_uv_dim, 
         main = '\nDiff: dim UV', ylab = 'diff kappamu',
         ylim = c(-2, 2), col = adjustcolor('purple', alpha.f = 100/255),
         axes = FALSE, cex.axis = 0.7)
axis(1); axis(2, at = -2:2); abline(h = 0, lty = 2)

# Plot 4: Interaction Effect
VertHist(data = diff_interaction, 
         main = '\nDim UV vs dim green', ylab = 'diff kappamu',
         ylim = c(-2, 2), col = adjustcolor('gray40', alpha.f = 100/255),
         axes = FALSE, cex.axis = 0.7)
axis(1); axis(2, at = -2:2); abline(h = 0, lty = 2)
```
