---
title: "GUV_Analysis_summary"
output: html_document
date: "`r Sys.Date()`"
author: "James Foster"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
editor_options: 
  markdown: 
    wrap: sentence
---

## Details

> **AUTHOR**
>
> James Foster $2025~06~05$

> **MODIFIED**
>
> James Foster $2025~06~11$

> **DESCRIPTION**
>
> Loads the reorganised data from `colour_dance_reorg.csv`, saved from `GUV_organisedata.R`, and generates plots to show the how dances changed as a function of light stimulus.

> **INPUTS**
>
> `colour_dance_reorg.csv`, `Edrich1979_brightness-accuracy.csv`

> **OUTPUTS**
>
> Plotted results.

> **CHANGES**
>
> -   

> **REFERENCES**
>
> Edrich, W., Neumeyer, C. and von Helversen, O. (1979). “Anti-sun orientation” of bees with regard to a field of ultraviolet light. *J. Comp. Physiol.* 134, 151–157.
>
> Rossel, S. and Wehner, R. (1984). Celestial orientation in bees: the use of spectral cues. *J. Comp. Physiol. A* 155, 605–613.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up

First load the functions and packages required for this project. These can all be found in `GUV_functions.R`, within the same Github repository.

```{r Load functions and packages}
source('GUV_functions.R')
```

Then load the data that will be used. This includes our own dance data, reorganised by `GUV_organisedata.R` as `colour_dance_reorg.csv`, as well as data from Edrich *et al.* (1979) on the relationship between accuracy (mean vector length), wavelength and intensity. This was manually extracted from Fig. 2 and saved in `"Edrich1979_brightness-accuracy.csv`.

```{r Load data}
#Dance angle data
cd = read.table(file = '1Data/colour_dance_reorg.csv', 
                header = T, 
                sep  = ',')
#Accuracy data
ed = read.table(file = '1Data/Edrich1979_brightness-accuracy.csv', 
                header = T, 
                sep  = ',')
#Inspect each dataset
print(cd)
print(ed)
```

## Format data

Ensure that each variable has the correct data format for subsequent analysis.

```{r Data formatting}
cd = within(cd,
            {
              ID = as.factor(ID) # beedance identifier as a factor
              date = as.factor(date) # date as a factor
              signed_angle = Mod360.180(bearing)  # bearing between -180 and 180
              angle = circular(rad(signed_angle),# bearing between -pi and pi
                               rotation = 'clock') # circular format suppresses later warnings
            }
)
u_id = with(cd, unique(ID)) # unique beedances
length(u_id)#169 beedances
```

Identify dances in which all four stimulus conditions were used.

```{r full condition}
luc = sapply(u_id,
              FUN = IndCond,
              dt = cd)

#Most individuals that made it to bright green 
full_ids = u_id[luc == 4]
length(full_ids)#19 individuals
#extract just those individuals

#find data for full condition individuals
full_cd = subset(cd, ID %in% full_ids)
```

## Plot the raw dance data

Plot all 169 dances

```{r Raw dances, echo=FALSE, warning=FALSE}
#TODO make function
#set up a sequence for the axes
xc = seq(from = -pi, to = pi-1e-16, length.out = 1e3)
par(mfrow = c(2, 2),
    mar = c(0,0,0,0))
par(pty = 's')

#add legend to 1st page
plot(x = NULL,
     xlim = c(-1,1),
     ylim = c(-1,1),
     pch = 19,
     axes = FALSE,
     xlab = '',
     ylab = '',
     main = ''
)
legend(x = 'center',
       legend = c('Green Bright',
                  'Green Dim',
                  'UV Bright',
                  'UV Dim',
                  'Sun azimuth (North = Up)'),
       col = c('green',
               'darkgreen',
               'magenta',
               'purple',
               'orange'),
       pch = c(21,21,21,21,NA),
       lty = c(NA, NA, NA, NA, 1),
       lwd = c(2,2,2,2,3)
       )
#loop through individuals
for(ii in u_id)
{
  mnv_gh = Mvec(subset(cd, ID %in% ii & colour %in% 'g' & brightn %in% 'h')$angle)
  mnv_gl = Mvec(subset(cd, ID %in% ii & colour %in% 'g' & brightn %in% 'l')$angle)
  mnv_uh = Mvec(subset(cd, ID %in% ii & colour %in% 'u' & brightn %in% 'h')$angle)
  mnv_ul = Mvec(subset(cd, ID %in% ii & colour %in% 'u' & brightn %in% 'l')$angle)
  plot(x = NULL,
       xlim = 100*c(-1,1),
       ylim = 100*c(-1,1),
       pch = 19,
       axes = FALSE,
       xlab = '',
       ylab = '',
       main = ''
       )
  abline(a = 0, b = 1, col = 'gray90')
  abline(a = 0, b = -1, col = 'gray90')
  abline(h = 0, v = 0, col = 'gray75')
  lines(x = 20*sin(xc), y = 20*cos(xc), lty = 3)
  lines(x = 40*sin(xc), y = 40*cos(xc), lty = 3)
  lines(x = 60*sin(xc), y = 60*cos(xc), lty = 3)
  lines(x = 80*sin(xc), y = 80*cos(xc), lty = 3)
  text(x = 1:4*20,
       y = c(0,0,0,0),
       labels = paste(' run', 1:4*20), 
       cex = 0.3,
       adj = c(0,1))
  with(subset(cd, ID %in% ii),
       {
  lines(x = c(0, 100*sin(sun_az_rad)),
        y = c(0, 100*cos(sun_az_rad)),
        col = adjustcolor('orange',alpha.f = 0.5),
        lwd = 2)
  points(x = run*sin(as.numeric(angle)),
         y = run*cos(as.numeric(angle)),
         bg = gray(level = 1.0,
                   alpha =  0.4),
         col = c('green', 'darkgreen', 'magenta', 'purple')
                [ifelse(colour %in% 'g',
                        yes = ifelse(brightn %in% 'h', yes = 1, no = 2),
                        no = ifelse(brightn %in% 'h', yes = 3, no = 4))],
        pch = 21,
        lwd = 2
       )
       }
  )
  lines(x = c(0,100*sin(mnv_gh['mu'])*mnv_gh['rho']), 
        y = c(0,100*cos(mnv_gh['mu'])*mnv_gh['rho']), 
        col = 'green', 
        lwd = 1)
  lines(x = c(0,100*sin(mnv_gl['mu'])*mnv_gl['rho']), 
        y = c(0,100*cos(mnv_gl['mu'])*mnv_gl['rho']), 
        col = 'darkgreen', 
        lwd = 1)
  lines(x = c(0,100*sin(mnv_uh['mu'])*mnv_uh['rho']), 
        y = c(0,100*cos(mnv_uh['mu'])*mnv_uh['rho']), 
        col = 'magenta', 
        lwd = 1)
  lines(x = c(0,100*sin(mnv_ul['mu'])*mnv_ul['rho']), 
        y = c(0,100*cos(mnv_ul['mu'])*mnv_ul['rho']), 
        col = 'purple', 
        lwd = 1)
  mtext(ii,side = 3,line = -2)
}
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Accuracy effects

Calculate the mean vectors for each set of waggle-runs under each lighting condition.

```{r Summarise, echo=FALSE}
#calculate mean vectors
mean_vectors = aggregate(angle~ID*brightn*colour+sun_az, # N.B. Including sun azimuth drops some cases without a time stamp
                         data = cd,
                         FUN = rho.circular
)
#correct names
mean_vectors = within(mean_vectors,
                      {mean_vector = angle; rm(angle)} # anlge now indicates a mean vector, not an angle
)
#add kappa
mle_estimates = aggregate(angle~ID*brightn*colour+sun_az,
                          data = cd,
                          FUN = MLE_est
)
#add to the summary table and 
#calculate inverse softplus kappa
mean_vectors = within(mean_vectors,
                      {
                        mu = deg(mle_estimates$angle[,'mu'])
                        kappa = mle_estimates$angle[,'kappa']
                        iskappa  = inv_softplus(kappa)
                      }
)
#find the full condition individuals
mean_vectors_full = subset(mean_vectors,
                           subset = ID %in% full_ids)
```

### Effect of brightness on mean vector length

#### Entire dataset

The first thing we notice is that in the low intensity conditions the accuracy is generally lower than for the high intensity conditions.

```{r Plot mean vectors, echo=FALSE}
boxplot(mean_vector~colour*brightn,
        data = mean_vectors,
        ylim = c(0,1),
        col = adjustcolor(c('green2',
                            'purple',
                            'darkgreen',
                            'purple4'), alpha.f = 0.5),
        outline = FALSE)
stripchart(mean_vector~colour*brightn,
           data = mean_vectors,
           add = TRUE,
           vertical = TRUE,
           method = 'jitter',
           col = adjustcolor(c('green2',
                               'purple',
                               'darkgreen',
                               'purple4'), alpha.f = 0.5),
           bg = gray(level = 0,
                     alpha = 0.1),
           pch = 21)
#Identify stimuli
stim = apply(X = expand.grid(c = c('g','u'),b = c('h','l')),
             FUN = paste,
             collapse = '',
             MARGIN = 1)
#Plot lines linking dances by the same individual
invisible(
  {
    lapply(X = u_id,
           FUN = Plt_mvec
           )
  })
abline(h = c(0,1))
```

#### Within individuals (balanced cases only)

This effect is even clearer if we look at just the full condition dances.

```{r Plot full condition mean vectors}
stripchart(mean_vector~colour*brightn,
           data = mean_vectors_full,
           # add = TRUE,
           vertical = TRUE,
           method = 'jitter',
           jitter = 0.02,
           col = adjustcolor(c('green2',
                               'purple',
                               'darkgreen',
                               'purple4'), alpha.f = 0.5),
           bg = gray(level = 0,
                     alpha = 0.1),
           pch = 21)
#Plot lines linking dances by the same individual
invisible(
  {
    lapply(X = full_ids,
           FUN = Plt_mvec
           )
  })
abline(h = c(0,1))
```

### Comparison with Edrich *et al.* (1979)

How does this compare with previous experiments in which a light spot (x°) was presented through bandpass filters and ND filters to produce different intensities? In that data, there was a clear sigmoidal relationship between intensity and accuracy (mean vector length).

#### Plot raw data from Edrich *et al.* (1979) Fig. 2

```{r Edrich raw data, echo=FALSE}
with(ed,
     {
     plot(x = log10(intensity),
          y = accuracy,
          xlim = c(9,14),
          ylim = c(0,1),
          main = 'Data from Edrich et al. 1979, Fig. 2',
            xlab = 'log10(intensity) (photons/cm2/s)',
            ylab = 'accuracy (r)',
          pch = 21,
          cex = 1.5,
          col = 'black',
          bg = sapply(paste(wavelength),
                       FUN = switch,
                       `354` = 'purple',
                       `429` = 'blue',
                       `535` = 'darkgreen',
                       'red')
          )
     abline(h = c(0,1))
     }
     )

legend(x = 'bottomright',
       legend = paste(sort(unique(ed$wavelength),
                     decreasing = TRUE), 'nm'),
       col = c('darkgreen','blue', 'purple'),
       pch = c(20, 20, 20))
```

This is best modelled with a nonlinear (psychometric model) following the beta distribution (for variables bounded between zero and one).

#### Reorganise data

```{r Prepare data for model fitting}
ed = within(ed,
            {
              logit_accuracy = qlogis(accuracy)
              log10_intensity = log10(intensity)
              colour = sapply(paste(wavelength),
                              FUN = switch,
                              `354` = 'UV',
                              `429` = 'blue',
                              `535` = '0green',
                              NA)
            }
            )
```

#### Fit psychometric curves

For each light stimulus, fit a sigmoid curve bounded between baseline accuracy (shared across stimuli) and maximum accuracy (`1 - lapses`). Inflection point and 80%-rise-region width vary between stimuli, maximising posterior probability for a beta distribution in `(0,1)`. To do this, we first define a nonlinear formula in `brms`.

```{r Set up formula}

#set up model fit
formula_beta = bf(
  formula = accuracy ~ 
    inv_logit(LogitBase) + (1 - inv_logit(LogitLapse) - inv_logit(LogitBase) ) *#curve region
    inv_logit( 4.39*(log10_intensity - Inflex) / exp(LogWidth) ) , #inflection-width curve
  # for each of these parameters, we can set up a separate formula 
  # that describes how to predict them from the data 
  #Base rate of correct choices: "
  LogitBase ~ 1, #Base rate of correct choices: "~ 1" gives the instruction "estimate the mean across all data"
  #Lapse rate on a log(odds) scale:
  LogitLapse ~ colour, #this is similar to the formula in our LMM example
  #inflection point of the initial curve:
  Inflex ~ colour, #N.B. this is similar to the intercept, so it does not include effects of stimulus level
  #log 80% width of the curve:
  LogWidth ~ colour, #N.B. this is similar to the slope, so all of its effects depend on stimulus level
  family = Beta(link = "identity", link_phi = 'log'),
  nl = TRUE)#the joint distribution for these parameters is undefined, and therefore the parameters themselves are "nonlinear"

```

Since this is a nonlinear model (there are terms both inside and outside the logistic transform) it is important to establish prior distributions for all parameters. These ensure that parameter estimates explore credible ranges during fitting. The baseline is expected to be around `r = 0.27`, while the maximum accuracy should be around `r = 0.95`, and their estimates should not converge (that would flatten the curve, or reverse its direction). The inflection points are expected around $10^{11}$ photons cm$^{-2}$ s$^{-1}$, with a rise-region width of $10^{3}$ photons cm$^{-2}$ s$^{-1}$. All colour effects may vary by a few orders of magnitude.

```{r}

#set up priors
prior_beta = get_prior(formula = formula_beta,
                     data = ed)
#set the prior distribution
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogitBase' &
                      coef %in% 'Intercept' 
                  ] = 'normal(-1,1)' #a normal distribution centred on plogis(-1) = 0.27
                  })
#this prior is automatically bounded between 0 and 1, 
#Lapse rate priors ---------------------------------------------------
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogitLapse' &
                      coef %in% 'Intercept' 
                  ] = 'normal(-3,2)' #a normal distribution centred on -3
                  })

#for all other fixed effects on lapse rate, we'll suggest values around 0 (no effect)
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogitLapse' &
                      coef %in% 'colourblue' 
                  ] = 'normal(0,2)' #a normal distribution:mean 0, sd 3
                  })
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogitLapse' &
                      coef %in% 'colourUV' 
                  ] = 'normal(0,2)' #a normal distribution:mean 0, sd 3
                  })

#Inflection point priors ---------------------------------------------
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'Inflex' &
                      coef %in% 'Intercept'
                  ] = 'normal(11,2)' #a normal distribution:mean 11, sd 2
                  })
#for all coefficients, we'll suggest values around 0 (no effect)
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'Inflex' &
                      coef %in% 'colourblue' 
                  ] = 'normal(0,2)' #a normal distribution:mean 0, sd 2
                  })

prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'Inflex' &
                      coef %in% 'colourUV' 
                  ] = 'normal(0,2)' #a normal distribution:mean 0, sd 2
                  })

#Rise region width priors --------------------------------------------
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogWidth' &
                      coef %in% 'Intercept' 
                  ] = 'normal(1,2)' #a normal distribution:mean 1, sd 3
                  })
#for all coefficients, we'll suggest values around 0 (no effect)
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogWidth' &
                      coef %in% 'colourblue'
                  ] = 'normal(0,2)' #a normal distribution:mean 0, sd 3
                  })
prior_beta = within(prior_beta, 
                  { prior[
                    class %in% 'b' & #just the fixed effects
                      nlpar %in% 'LogWidth' &
                      coef %in% 'colourUV'
                  ] = 'normal(0,2)' #a normal distribution:mean 0, sd 3
                  })
```

Fit a psychometric model to the data using the nonlinear formula and specified prior distributions. `4000` iterations on `4` chains takes less than one minute.

```{r Fit model}
#takes less than 1 minute
system.time(
  {
    beta_fit = brm( formula = formula_beta, # using our nonlinear formula
                     data = ed, # our data
                     prior = prior_beta, # our priors 
                     iter = 4000, # long run for 4000 iterations
                     chains = 4, # 4 chains in parallel
                     cores = 4, # on 4 CPUs
                     refresh = 0, # don't echo chain progress
                     control = list(adapt_delta = 0.95), #finer sampling
                     backend = 'cmdstanr') # use cmdstanr (other compilers broken)
  }
)
```

#### Plot and compare data from Edrich and Kühn.

Summarise the data from full cases in the Kühn data using quantiles (minimum, 25%, median, 75% and maximum).

```{r Collect relevant data from Kühn data}
#Examine quantiles
meanQ = aggregate(mean_vector ~ colour*brightn, 
                  data = mean_vectors_full,
                  FUN = quantile)
#Extract different quantiles as separate columns in a data frame
meanQ = data.frame(cbind(meanQ[1:2], meanQ$mean_vector) )
#Rename
names(meanQ) = c('colour', 'brightn', paste0('q',0:4))
#Subset by condition
gh_meanQ = subset(meanQ, colour %in% 'g' & brightn %in% 'h')
gl_meanQ = subset(meanQ, colour %in% 'g' & brightn %in% 'l')
uh_meanQ = subset(meanQ, colour %in% 'u' & brightn %in% 'h')
ul_meanQ = subset(meanQ, colour %in% 'u' & brightn %in% 'l')

```

Extract the model predictions by summarising parameter estimates from the sampling phases of all chains. To identify specific combinations of parameter values, use the `conditional_effects` function, which organises all simultaneous estimates into a data frame.

```{r Extract model predictions}

# #N.B. for plotting we will need to also define "inv_logit" in our the R environment
inv_logit = inv_logit_scaled # this can be found in brms, but with an additional specifier

#summary of parameter estimates
full_sm_beta = summary(beta_fit,
                  robust = TRUE)#use the median estimate

#Extract fixed effects estimates
full_fix_beta = full_sm_beta$fixed
#extract rownames
full_fix_rn_beta = rownames(full_fix_beta)

# stimulus levels to add to the dataset to be able to measure this effect.
full_estimates_beta = full_fix_beta$Estimate
names(full_estimates_beta) = full_fix_rn_beta

# to predict all effects, we can use the 'posterior_epred' method
#by default 100 predictions per continuous variable (but fewer for their interactions)
#This takes <2 seconds
system.time(
  {
    full_cond_beta =brms::conditional_effects(beta_fit, 
                                         method = 'posterior_epred', # posterior epred not working
                                         cores =  parallel::detectCores()-1,
                                         effects = c('log10_intensity:colour'),
                                         plot = FALSE
    )
  }
)
#extract predictions
pred_data_beta = full_cond_beta$`log10_intensity:colour`

#Green light curve
g_curve = 
with(data.frame(t(full_estimates_beta)),
     {
       list(Base = plogis(LogitBase_Intercept),
            Maxima = 1- plogis(LogitLapse_Intercept), 
            Inflection = Inflex_Intercept,  
            Width = exp(LogWidth_Intercept)  
       )
     }
)

#UV light curve
u_curve = 
with(data.frame(t(full_estimates_beta)),
     {
       list(Base = plogis(LogitBase_Intercept),
            Maxima = 1- plogis(LogitLapse_Intercept + LogitLapse_colourUV), 
            Inflection = Inflex_Intercept + Inflex_colourUV,  
            Width = exp(LogWidth_Intercept + LogWidth_colourUV)  
       )
     }
)



```

Plot the raw data alongside the predictions for the data from Edrich *et al*., (1979) and a boxplot summary of the data data from Kühn (2017).

```{r Plot model predictions}

#Prepare a sequence for plotting
xxl = seq(from = 9, to = 16,
          length.out = 1e3)

#plot only the stimulus types also used by Kühn.
ed_ug = subset(ed, wavelength != 429)

#plot each stimulus type
with(ed_ug,
     {
       plot(x = log10_intensity,
            y = accuracy,
            main = 
'Comparison of data from Edrich et al. 1979
with Kühn, 2017',
            xlab = 'log10(intensity) (photons/cm2/s)',
            ylab = 'accuracy (r)',
            ylim = c(0,1),
            xlim = c(9,16),
            pch = 21,
            cex = 1.5,
            col = 'black',
            bg = sapply(paste(wavelength),
                        FUN = switch,
                        `354` = 'purple',
                        `429` = 'blue',
                        `535` = 'darkgreen',
                        'red')
       )
       abline(h = c(0,1))
       abline(v = log10(4*c(1e12,1e14)) )
     }
)
#Add the estimated range of the curve for the green stimulus
with(data.frame(t(full_estimates_beta)),
     {
  abline(h = c(0,1,
               plogis(LogitBase_Intercept),
               1- plogis(LogitLapse_Intercept) ), 
         lty = c(1,1,3,3)
         )
     }
)

#plot total 95% prediction intervals
with(subset(pred_data_beta, colour == '0green'), 
     {
       polygon(x = c(sort(log10_intensity), rev(sort(log10_intensity))), 
               y = c(lower__[order(log10_intensity)],
                     rev(upper__[order(log10_intensity)])
               ), 
               col = adjustcolor('green', alpha.f = 25/256),
               border = NA,
               lwd = 0.1
       )
     }
)
with(subset(pred_data_beta, colour == 'UV'), 
     {
       polygon(x = c(sort(log10_intensity), rev(sort(log10_intensity))), 
               y = c(lower__[order(log10_intensity)],
                     rev(upper__[order(log10_intensity)])
               ), 
               col = adjustcolor('magenta', alpha.f = 25/256),
               border = NA,
               lwd = 0.1
       )
     }
)

#plot the median prediction lines
with(g_curve, 
     lines(x = xxl,
           y = Base + (Maxima  - Base) * 
             plogis(4.39*(xxl - Inflection)/Width), 
           col = 'darkgreen',
           lwd = 3,
           lty = 3)
)
with(u_curve, 
     lines(x = xxl,
           y = Base + (Maxima  - Base) * 
             plogis(4.39*(xxl - Inflection)/Width), 
           col = 'purple4',
           lwd = 3,
           lty = 3)
)

#plot observed data from Kühn as boxplots showing the quartiles, median and total range.
#Green High intensity
QBox(quant = subset(meanQ, colour %in% 'g' & brightn %in% 'h'),
     x = log10(4e14),
     offset = -0.1,
     cols = c('green', 'darkgreen')
    )

#Green Low intensity
QBox(quant = subset(meanQ, colour %in% 'g' & brightn %in% 'l'),
     x = log10(4e12),
     offset = -0.1,
     cols = c('green', 'darkgreen')
    )

#UV High intensity
QBox(quant = subset(meanQ, colour %in% 'u' & brightn %in% 'h'),
     x = log10(4e14),
     offset = 0.1,
     cols = c('magenta', 'purple')
    )

#UV Low intensity
QBox(quant = subset(meanQ, colour %in% 'u' & brightn %in% 'l'),
     x = log10(4e12),
     offset = 0.1,
     cols = c('magenta', 'purple')
    )

legend(x = 'bottomright',
       legend = c(
                   paste('Edrich', 
                        sort(unique(ed_ug$wavelength),
                           decreasing = TRUE), 'nm'),
                   paste('  Kühn', c(528, 365), 'nm' ) 
       ),
       col = c('darkgreen', 'purple', 'green', 'magenta'),
       pch = c(20, 20, 15, 15),
       lty = c(3,3,NA, NA),
       lwd = 3,
       cex = 0.7)
```

It would be difficult to explain the orientation accuracy for the high intensity ($4*10^{14}$) UV stimulus from the accuracy curve for Edrich's UV data, but it would fit well with the maximum accuracy observed for green light. In Edrich's study, the green light was interpreted as the sun (dances at correct angle between sun and feeder), whereas the UV light was interpreted as a patch of sky near the anti-sun. In Kühn's data the stimuli were smaller (assuming a viewing distance of ≈200mm, an SMD with diameter 3.5mm would appear `deg(2*atan2(y = 3.5/2,x = 200)) =` 1.0° across) than Edrich's (2.3°), which may have resulted in them being perceived as dimmer (left shift in this graph).

# Sun compass orientation

If most dances are for the same location, we should observe dances as the same fixed angle to the stimulus, interpreted either as the sun or the anti-sun.

```{r Calculate and plot angles relative expected sun azimuth}

sun_diff_gh = sapply(X = u_id,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'g',
                    br = 'h')
sun_diff_gl = sapply(X = u_id,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'g',
                    br = 'l')
sun_diff_uh = sapply(X = u_id,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'u',
                    br = 'h')
sun_diff_ul = sapply(X = u_id,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'u',
                    br = 'l')


par(mfrow = c(2,2), mar = c(0,0,0,0))
PCfun(angles = unlist(sun_diff_gh),
      col = 'green',
      shrink = 2.0,
      side = 3,
      title = 'Green Bright')
PCfun(angles = unlist(sun_diff_gl),
      col = 'darkgreen',
      shrink = 2.0,
      side = 3,
      title = 'Green Dim')
PCfun(angles = unlist(sun_diff_uh),
      col = 'magenta',
      shrink = 2.0,
      side = 3,
      title = 'UV Bright')
PCfun(angles = unlist(sun_diff_ul),
      col = 'purple',
      shrink = 2.0,
      side = 3,
      title = '           UV Dim')
```

For the bright stimuli the dances generally indicate South, but with a very wide range. (or possibly clusters to the South-West and South-East). For the dim stimuli there is less of a clear pattern, particularly for UV dim.

We can also restrict this analysis to just the dances with all conditions.

```{r Angles relative expected sun azimuth for full conditions}

full_sun_diff_gh = sapply(X = full_ids,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'g',
                    br = 'h')
full_sun_diff_gl = sapply(X = full_ids,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'g',
                    br = 'l')
full_sun_diff_uh = sapply(X = full_ids,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'u',
                    br = 'h')
full_sun_diff_ul = sapply(X = full_ids,
                    FUN = SunDiff,
                    dt = mean_vectors,
                    cl = 'u',
                    br = 'l')


par(mfrow = c(2,2), mar = c(0,0,0,0))
PCfun(angles = unlist(full_sun_diff_gh),
      col = 'green',
      shrink = 1.5,
      side = 3,
      title = 'Green Bright')
PCfun(angles = unlist(full_sun_diff_gl),
      col = 'darkgreen',
      shrink = 1.5,
      side = 3,
      title = 'Green Dim')
PCfun(angles = unlist(full_sun_diff_uh),
      col = 'magenta',
      shrink = 1.5,
      side = 3,
      title = 'UV Bright')
PCfun(angles = unlist(full_sun_diff_ul),
      col = 'purple',
      shrink = 1.5,
      side = 3,
      title = 'UV Dim')
      
```

When restricting to only the balanced cases, the pattern becomes clearer. For the most oriented dances to the bright UV stimulus there are two clusters to the South-West and South-East. These are likely also present for the bright green stimulus. The dim green stimulus appears to align somewhat with the bright stimuli. The dim UV stimulus shows no apparent patten, though a slight southwards bias.
